---
title: "Hands-on Exercise 10"
author: Johsuan Huang
date: 2024-03-27
date-modified: "last-modified"
description: "Information Dashboard Design: R methods"
categories: [ggthemes,reactable,reactablefmtr, gt, gtExtras, tidyverse]
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

In this hands-on exercise, we will be able to:

-   create bullet chart by using ggplot2

-   create sparklines by using ggplot2

-   build industry standard dashboard by using R Shiny

# 2 Getting Started

## Load the packages

-   [gtExtras](https://jthomasmock.github.io/gtExtras/index.html) provides some additional helper functions to assist in creating beautiful tables with [gt](https://gt.rstudio.com/), an R package specially designed for anyone to make wonderful-looking tables using the R programming language.

-   [reactable](https://glin.github.io/reactable/index.html) provides functions to create interactive data tables for R, based on the [React Table](https://github.com/tanstack/table/tree/v7) library and made with [reactR](https://github.com/react-R/reactR).

-   [reactablefmtr](https://kcuilla.github.io/reactablefmtr/index.html) provides various features to streamline and enhance the styling of interactive reactable tables with easy-to-use and highly-customizable functions and themes.

```{r}
pacman::p_load(lubridate, ggthemes, reactable,
reactablefmtr, gt, gtExtras, tidyverse,odbc,svglite)
```

# 3 Import the data

For the purpose of this study, a personal database in Microsoft Access mdb format called *Coffee Chain* will be used:

Microsoft Access databases (.mdb, .accdb) are not natively supported on macOS because there is no official Microsoft Access ODBC driver for macOS. Instead, I will use `read_csv()` with `pipe()` to extract data directly from the database and load it into R:

-   `system()`: Run a system command from within R

-   `pipe()`: Run a shell Command from R and read command output directly into R

-   `read_csv()`: Reading the Exported Data

We first check tables' name in the mbd file:

```{r}
# check table names in the mdb file
system('mdb-tables "data/Coffee Chain.mdb"')

```

Next, we import three tables:

```{r}

# Read each table
factTable <- read_csv(pipe('mdb-export "data/Coffee Chain.mdb" "factTable"'))
location <- read_csv(pipe('mdb-export "data/Coffee Chain.mdb" "location"'))
product <- read_csv(pipe('mdb-export "data/Coffee Chain.mdb" "product"'))

```

After analyzing the data, we merged the fact table with the location and product tables using a left join:

```{r}

coffeechain <- factTable %>%
  left_join(location, by = "Area Code") %>%
  left_join(product, by = "ProductId")
```

## 3.1 Check data structure and data summary

The data format of "Date" is "chr". We should transform the variable into date format.

```{r}
str(coffeechain)
```

```{r}
coffeechain <- coffeechain %>%
  mutate(Date = as.Date(Date,format="%m/%d/%y"))
```

```{r}
summary(coffeechain)
```

## 3.2 Data Preparation

The code chunk below is used to aggregate Sales and Budgeted Sales at the Product level:

```{r}
product <- coffeechain %>% 
  group_by(Product) %>%
  summarise(target = sum(`Budget Sales`),
            current = sum(Sales)) %>%
  ungroup()
```

# 4 Plotting Bullet chart & sparklines using ggplot2

## 4.1 Plotting Bullet chart

The code chunk below is used to plot the bullet charts using ggplot2 functions.

The plot shows the sales target by product (red error bar) along with 75% and 85% of the target sales:

```{r}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

## 4.2 Plotting Sparklines

### 4.2.1 Prepare the data

The code chunk below is to aggregate sales by Month and product:

```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```

The code chunk below is used to compute the minimum, maximum and end othe the month sales:

-   `slice()`: selects rows from the grouped data.

-   `which.min(Sales)` finds the index of the minimum value in the Sales column for each product group.

```{r}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))
```

The code chunk below is used to compute the 25 and 75 quantiles:

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 0.25),
            quart2 = quantile(Sales, 0.75)) %>%
  right_join(sales_report)
```

### 4.2.2 sparklines in ggplot2

The sparkline plot displays monthly sales, highlighting maximum and minimum values, and shows sales in December by product.

The code below uses `geom_ribbon` to compare total sales amounts, where the higher the sales, the wider the ribbon:

```{r}
#| fig-height: 6
#| fig-width: 12

ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey80') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.35) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  labs(title= "Monthly Sales Breakdown by Product")+
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(plot.title = element_text(size=15,hjust=0.5,face="bold"),
        axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank(),
        plot.background=element_rect(fill="#f3f1e9",colour="#f3f1e9"))
```

# 5 Static Information Dashboard Design: gt and gtExtras methods

In this section, we will learn how to create static information dashboard by using [gt](https://gt.rstudio.com/) and [gtExtras](https://jthomasmock.github.io/gtExtras/index.html) packages.

## 5.1 Plotting a simple bullet chart

```{r}
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("#A4A9D9", 
                          "black")) %>%
gt_theme_espn()
```

## 5.2 sparklines: gtExtras method

Before we can prepare the sales report by product by using gtExtras functions, code chunk below will be used to prepare the data.

```{r}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales),.groups = "drop") %>%
  ungroup()
```

It is important to note that one of the requirement of gtExtras functions is that almost exclusively they require you to pass **data.frame with list columns**. In view of this, code chunk below will be used to convert the report data.frame into list columns.

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

# 6 Plotting Coffechain Sales report

## 6.1 Plotting Sparklines

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE) %>%
  tab_style(
    style = cell_fill(color = "#f3f1e9"), 
    locations = cells_body()
  ) %>%
  tab_style(
    style = list(cell_fill(color = "#f3f1e9"),
                 cell_text(weight = "bold")),
    locations = cells_column_labels())

```

## 6.2 Adding statistics

First, calculate summary statistics by using the code chunk below.

```{r}
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = c(2,3,4),
    decimals = 0)%>%
  gt_theme_espn()
```

## 6.3 Combining the data.frame

Next, use the code chunk below to add the statistics on the table.

```{r}
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")

sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )

sales_data = left_join(sales, spark)

```

## 6.4 Plotting the updated data.table

```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE) %>%
    fmt_number(columns = c(2,3,4),
    decimals = 0)%>%
    gt_theme_espn()
```

## 6.5 Combining bullet chart and sparklines

```{r}
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup() 

sales_data <- sales_data %>%
  left_join(bullet)

sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("#A4A9D9", 
                          "black")) %>%
      fmt_number(columns = c(2,3,4),
    decimals = 0)%>%
  gt_theme_538()
```

# 7 Interactive Information Dashboard Design: reactable and reactablefmtr methods

In this section, we will learn how to create interactive information dashboard by using [reactable](https://glin.github.io/reactable/index.html) and [reactablefmtr](https://kcuilla.github.io/reactablefmtr/index.html) packages.

In order to build an interactive sparklines, we need to install dataui R package by using the code chunk below.

```{r}
remotes::install_github("timelyportfolio/dataui")
```

```{r}
library(dataui)
```

## 7.1 Plotting interactive sparklines

Similar to gtExtras, to plot an interactive sparklines by using reactablefmtr package we need to prepare the list field by using the code chunk below.

```{r}
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))
```

Next, [react_sparkline](https://kcuilla.github.io/reactablefmtr/reference/react_sparkline.html) will be to plot the sparklines as shown below.

```{r}
reactable(
  report,
  defaultPageSize = 13, #By default the pagesize is 10
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

## 7.2 Adding points and labels

In the code chunk below `highlight_points` argument is used to show the minimum and maximum values points and `label` argument is used to label first and last values.

```{r}
#| fig-width: 12

reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200,
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9")),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        ),
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9"),
    )
  )
)
```

::: {.callout-note appearance="simple"}
-   **ggraph** introduces a special ggplot theme that provides better defaults for network graphs than the normal ggplot defaults. `theme_graph()`, besides removing axes, grids, and border, changes the font to Arial Narrow (this can be overridden).

-   The ggraph theme can be set for a series of plots with the `set_graph_style()`command run before the graphs are plotted or by using `theme_graph()` in the individual plots.
:::

## 7.3 Adding reference line

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(
      maxWidth = 200,
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9")
      ),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        ),
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9"),
    )
  )
)

```

## 7.4 Adding bandline

Instead adding reference line, bandline can be added by using the `bandline` argument.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(
      maxWidth = 200,
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9")
      ),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        ),
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9"),
    )
  )
)
```

## 7.5 Changing from sparkline to sparkbar

Instead of displaying the values as sparklines, we can display them as sparkbars as shiwn below.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(
      maxWidth = 200,
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9")
      ),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean"
        ),
      style = list(backgroundColor = "#f3f1e9"),
      headerStyle = list(backgroundColor = "#f3f1e9"),
    )
  )
)
```

# 8 Reference

-   Kam, T.S. (2025). [Information Dashboard Design: R methods](https://r4va.netlify.app/chap31)
