---
title: "Take-Home Exercise 2: Be Tradewise or Otherwise"
author: Johsuan Huang
date: 2024-02-24
date-modified: "last-modified"
categories: []
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

note:

-   The website should add explanations for the relationships between merchandise trade, imports and exports (domestic export + re-export), and highlight the difference between diff export type and export of service

-   <https://www.singstat.gov.sg/modules/infographics/singapore-international-trade>

makeover:

1.  **MERCHANDISE TRADE PERFORMANCE WITH MAJOR TRADING PARTNERS, 2024**

    The plot clearly shows the net export deficit and surplus partners of Singapore. However, it would be more interesting and informative if we can make it into an animation.

2.  **Non-oil re-exports**

    The data is organized in the shape of boats, which is creative and attractive. However, due to ships being displayed vertically and sparse data points, it's difficult to read the information. Additionally, it only shows data from 2024, so we cannot see the variations between years. **May use newggslopegraph to show the trend.**

3.  **Import of Service**

    It uses different icons and colors to highlight different service sectors, which make it easier for users to recognize the import amount and percentage by service sectors from the graph. However, since the number of service sectors is 12, it is a bit messy to display the information by pie chart.

    **May consider:**

    -   Combine the service industries with lower proportions (5%), such as insurance and finance
    -   Use tree map and animation to show the variations between years and sectors.

## 1 Overview

Since Mr. Donald Trump took office as the President of the United States on January 20, 2025, one of the most closely watched topics has been global trade.

As a visual analytics novice, you are eager to apply newly acquired techniques to explore and analyze the changing trends and patterns of Singapore’s international trade since 2015.

## 2 Getting Started

::: panel-tabset
## Load the packages and import data

Use the **pacman** package `p_load()` to check, install and launch the following R packages:

```{r}
pacman::p_load(tidyverse, lubridate, gridExtra,
               CGPfunctions,patchwork,ggthemes,gganimate,
               ggtext,scales,patchwork,ggrepel,
               ggridges, ggdist, ggiraph, plotly, 
               knitr,ggiraph,ggstatsplot,readxl,
               treemap
               )
```

| **Library** | **Description** |
|-------------------------|------------------------------------------------|
|  |  |
| tidyverse | a family of R packages for data processing |
| [**Plotly R**](https://plot.ly/r/) | an R package for creating interactive web-based graphs via plotly’s JavaScript graphing library. |

Use read_excel function to import xlsx files:

```{r}
m_import <- read_excel("data/SGTrade_byCountry.xlsx",sheet = "T1",range="A11:BJ171")
m_export <- read_excel("data/SGTrade_byCountry.xlsx",sheet = "T2",range="A11:BJ171")
m_re_export <- read_excel("data/SGTrade_byCountry.xlsx",sheet = "T3",range="A11:BJ171")
```

## Observe data

```{r}
kable(head(m_import,10))
```

```{r}
#str(pop_data)
```

```{r}
#summary(pop_data)
```

```{r}
any(is.na(m_import))
any(is.na(m_export))
any(is.na(m_re_export))

```

## Prepare data

From the "Observe data" tab, we can see that age is a categorical variable. For better analysis, we should transpose this variable and group the ages into young, active, and old categories to make the data more interpretable:

There are 10 age groups in this dataset:

```{r}
#pop_data %>% distinct(AG)
```

```{r}
# Adding new derived variables for the young, economy active and old age group


```
:::

## 3 GOV Trade Data Visualization Make-Over

### **3.1 MERCHANDISE TRADE PERFORMANCE WITH MAJOR TRADING PARTNERS, 2024**

-   **Pros:** Use the diagonal line and blocks to clearly show **net export deficit and surplus partners** of Singapore

-   **Cons:** It only shows 2024 mechandise data, so it cannot offer a complete view of merchandise trade trends in recent years

-   **Make-Over Suggestion**: Use animation combining years of data to enhance data storytelling

::: panel-tabset
## Original

## Make-over

### 
:::

#### 3.1.1 Import data

```{r}
m_import <- read_excel("data/SGTrade_byCountry.xlsx",sheet = "T1",range="A11:DR171")
m_export <- read_excel("data/SGTrade_byCountry.xlsx",sheet = "T2",range="A11:DR171")
m_re_export <- read_excel("data/SGTrade_byCountry.xlsx",sheet = "T3",range="A11:DR171")
```

#### 3.1.2 Data processing

```{r}

# rename the col name of country
colnames(m_export)[1] <- "country"
colnames(m_re_export)[1] <- "country"
colnames(m_import)[1] <- "country"

# copy the structure and transform it to long df 
m_by_country <- m_export %>%
  gather(key = month_year, value="export",-country)

m_re_export_long <- m_re_export %>%
  gather(key = month_year, value="re_export",-country)

m_import_long <- m_import %>%
  gather(key = month_year, value = "import",-country)

# join all data together
m_by_country <- m_by_country %>%
  left_join(m_re_export_long,by=c("country","month_year")) %>%
  left_join(m_import_long, by = c("country","month_year")) %>%
  mutate(total_trade = round(export+re_export+import,1),
         net_export = round((export+re_export)-import,1),
         abs_net_export = abs(round((export+re_export)-import,1)),
         year = as.integer(substr(month_year,1,4)))%>%
  filter(year != 2025)

# create yearly data
m_by_country_year <- m_by_country %>%
  group_by(country,year) %>%
  summarise(across(c("export","re_export","import","net_export","abs_net_export","total_trade"),~sum(.x, na.rm = TRUE)/1000),.groups="drop")

# filter continent only
m_by_continent_year <- m_by_country_year %>%
  filter(country %in% c("America","Asia","Europe","Oceania","Africa","Total All Markets"))

# filter country only
m_by_country_year <- m_by_country_year %>%
  filter(!country %in% c("America","Asia","Oceania","Africa","Total All Markets"))

# calculate avg net export 
m_by_country_avg <- m_by_country_year %>%
  group_by(country)%>%
  summarise(across(c("export","re_export","import","net_export","abs_net_export","total_trade"),~round(mean(.x),1)),.groups="drop") %>%
  arrange(desc(total_trade))

# top 5 postive net export countries for SG
top_10_m_trade <- m_by_country_avg$country[1:10]

# filter top 10 trade countries
top_10_trade_country <- m_by_country_year %>%
  filter(country %in% top_10_m_trade) %>%
  mutate(total_export = re_export + export)

country_abbr <- c(
  "Hong Kong" = "HK",
  "Indonesia" = "ID",
  "Malaysia" = "MY",
  "Thailand" = "TH",
  "China" = "CN",
  "Taiwan" = "TW",
  "United States" = "US",
  "Europe" = "EU",
  "Korea, Rep Of" = "KR",
  "Japan" = "JP"
)

top_10_trade_country <- top_10_trade_country %>%
  mutate(country_abbr = country_abbr[country])

```

```{r}
summary(top_10_trade_country)
```

#### 3.1.3 Plotting

```{r}
#| echo: false
#| fig-height: 6
#| fig-width: 10
#| out-extra: "style='max-width:100%; display: block;'"
#| out-width: 100%

my_palette <- c("#D989AE", "#96C6D9", "#F2CD88", "#D97259", "#468C6C", "grey60", 
                      "#B3A2C7", "#E1A77C", "#7CA08B", "#C77C8C")

ggplot(data=top_10_trade_country,
       aes(x= total_export, y = import))+
  # lower part background
  annotate(geom = "polygon", 
           x = c(-Inf, Inf, Inf),
           y = c(-Inf, Inf, -Inf), 
           fill = "#34a853", alpha = 0.1)+
  # upper part background
  annotate(geom = "polygon", 
           x = c(Inf, -Inf, -Inf),
           y = c(Inf, -Inf, Inf), 
           fill = "#4285f4", alpha = 0.1)+
  annotate("text", x = 10, y = 90, label = "Net Export\nDeficit", 
           color = "#4285f4", size = 4, hjust = 0.5, alpha=0.7) +
  annotate("text", x = 90, y = 20, label = "Net Export\nSurplus", 
           color = "#34a853", size = 4, hjust = 0.5, alpha=0.7) +
  geom_point(
       aes(size = total_trade, color = country_abbr),
       alpha=0.7,
       show.legend = FALSE)+
         scale_colour_manual(values = my_palette)+
         scale_size(range = c(2,15))+
         labs(title = "Merchandise Trade Performance With Top 10 Trading Partners",
         subtitle = "Year: {frame_time}",
         x = "Annual Exports<br>S$ Bil",
         y = "Annual Imports<br>S$ Bil",
         caption = "Source: singstat.gov.sg ")+
  geom_label(aes(label = country_abbr,
                 color = country_abbr),
             hjust=0.5, 
             vjust=-1,
             fill = NA,
             label.size = 0,
             show.legend = FALSE)+
  transition_time(year)+
  ease_aes('cubic-in-out')+
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100))+
  theme_classic()+
  theme(
    plot.title = element_markdown(size = 15, hjust = 0,face="bold"),
    axis.line.x = element_line(color = "#34a853"),
    axis.ticks.x = element_line(color = "#34a853"),
    axis.ticks.y = element_line(color = "#4285f4"), 
    axis.line.y = element_line(color = "#4285f4"),  
    axis.text.x = element_text(color = "#34a853"), 
    axis.text.y = element_text(color = "#4285f4"),
    axis.title.x = element_markdown(color = "#34a853",
                                margin = margin(t = 13),
                                hjust=1), 
    axis.title.y = element_markdown(color = "#4285f4",
                                margin = margin(r = 10),
                                hjust=1 ,angle = 360),
    panel.background = element_rect(fill = "#f3f1e9"),
    plot.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.margin = margin(1, 1, 1, 1, "cm"), 
    panel.spacing = unit(2, "cm"))
  
```

### **3.2 Non-oil re-exports**

-   **Pros:** Creative and attractive visual design (stacked bar chart shaped like boats)

-   **Cons:**

    1.  Only shows 2024 data, making it hard to compare trends over time and less meaningful

    2.  Difficult to read due to vertical ship display and sparse data points

-   **Make-Over Suggestion**: Use **newggslopegraph** to show trends over multiple years

::: panel-tabset
## Original

## Make-Over
:::

#### 3.2.1 Import data and Data Processing

```{r}
com_re_export <- read_excel("data/SGTrade_byCommodity.xlsx",sheet = "T1",range="A11:DR79")

non_oil_re_export <- com_re_export[-(1:59),] 

colnames(non_oil_re_export)[1] = "commodity"

non_oil_re_export <- non_oil_re_export %>%
  gather(key="month_year",value="re_export",-commodity)%>%
  mutate(year = as.ordered(substr(month_year,1,4)))%>%
  filter(year %in% c(2015,2020,2022,2024))%>%
  group_by(commodity,year)%>%
  summarise(re_export = round(sum(re_export,na.rm=TRUE)/1000000),.groups = "drop")

```

### 3.2.2 Plotting

```{r}
#| echo: false
#| fig-height: 6
#| fig-width: 10
#| out-extra: "style='max-width:100%; display: block;'"
#| out-width: 100%



p1 <- non_oil_re_export %>%
  group_by(year) %>%
  mutate(total_re_export_year = sum(re_export, na.rm = TRUE), 
         proportion = round(re_export / total_re_export_year * 100),
         label = ifelse(year %in% c(2015,2020,2022,2024),
                        paste0("$ ",comma(re_export)," B / ",proportion,"%"),
                        NA))%>% 
  ungroup()%>%
  filter(commodity == "Machinery & Transport Equipment") %>%
  newggslopegraph(year, re_export, commodity,
                  Title = "<span style='color:grey30;'>NON-OIL RE-EXPORT by MAJOR COMMODITY</span><br>",
                  SubTitle = "**Something was brewing during the COVID-19 padenmic?**<br><span style='color:#D989AE;'>**Machinery & Transportation Equipment**</span> not only remained unaffected by COVID,<br>but saw a **37% surge** in re-export trade during the period.<br> <br> ",
                  Data.label = label,
                  Caption = NULL,
                  DataLabelPadding = 0.5,
                  DataLabelFillColor = "#f3f1e9",
                  LineColor = "#D989AE") +
  annotate("rect", xmin = 2, xmax = 3, ymin = -Inf, ymax = Inf, 
         fill = "gray", alpha = 0.25)+
    geom_segment(aes(x = 1.7, y = 240, xend = 2, yend = 250),
               linetype = "solid", color = "grey20", size = 0.5,
               arrow = arrow(length = unit(0.03, "npc"))) +
    geom_richtext(
    aes(x = 1.7, y = 240, 
        label = "2020–2022 was the<br>**COVID-19 pandemic**"),
    fill = NA, label.color = NA, 
    color = "grey40", size = 3, hjust = 1, alpha = 0.7
  ) +
  theme(axis.ticks = element_blank(),
        plot.title = element_markdown(size = 15, hjust = 0.5, face = "bold",lineheight = 0.8),
        plot.subtitle = element_markdown(size = 12, hjust = 0.5,lineheight = 1.3),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        panel.background = element_rect(fill = "#f3f1e9", color = NA),
        plot.background = element_rect(fill = "#f3f1e9", color = NA),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())


p2 <- non_oil_re_export %>%
  group_by(year) %>%
  mutate(total_re_export_year = sum(re_export, na.rm = TRUE), 
         proportion = round(re_export / total_re_export_year * 100),
         label = ifelse(year %in% c(2015,2024),
                        paste0("$ ",comma(re_export)," B / ",proportion,"%"),
                        NA))%>% 
  ungroup()%>%
  filter(commodity != "Machinery & Transport Equipment") %>%
  
  newggslopegraph(year, re_export, commodity,
                  Title = NULL,
                  SubTitle = NULL,
                  Caption = "Source: singstat.gov.sg",
                  Data.label = label,
                  DataLabelPadding = 0,
                  DataLabelFillColor = "#f3f1e9",
                  LineColor = c(rep("#B0B0B0",8))) + 
  scale_x_discrete(position = "bottom")+
  annotate("rect", xmin = 2, xmax = 3, ymin = -Inf, ymax = Inf, 
         fill = "gray", alpha = 0.25)+
  theme(axis.ticks = element_blank(),
        plot.title = element_text(size = 10, hjust = 0, face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        panel.background = element_rect(fill = "#f3f1e9", color = NA),
        plot.background = element_rect(fill = "#f3f1e9", color = NA),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())



p1 <- p1 + theme(plot.margin = margin(20, 5, -5, 5))
p2 <- p2 + theme(plot.margin = margin(-5, 5, 10, 5))
combined_plot <- p1 / p2
combined_plot

```

### **3.3 Import of Service**

-   **Pros**:

    1.  Icons and colors help differentiate service sectors, making it visually appealing

    2.  Allows users to easily recognize import amounts and percentages

-   **Cons**:

    1.  Too cluttered (12 service sectors in a pie chart)

    2.  Hard to compare variations across years

-   **Make-Over Suggestion**:

    1.  Use a **tree map** to improve readability.

    2.  Add **animation** to show changes over years and across sectors.

```{r}
s_export <- read_excel("data/SGTrade_byService.xlsx",sheet = "T2",range="A11:k55")

service_type <- c(
  "Manufacturing Services On Physical Inputs Owned By Others",
  "Maintenance And Repair Services",
  "Transport",
  "Travel",
  "Insurance",
  "Government Goods And Services",
  "Construction",
  "Financial",
  "Telecommunications, Computer & Information",
  "Charges For The Use Of Intellectual Property",
  "Personal, Cultural And Recreational",
  "Other Business Services",
  "Advertising And Market Research",
  "Business Management"
)

colnames(s_export)[1]="service"

s_export_selected <- s_export %>%
  filter(service %in% service_type)

# deduct Ad & business management from other Business service 

s_export_selected[12,2:ncol(s_export_selected)] <- s_export_selected[12,2:ncol(s_export_selected)] - s_export_selected[13,2:ncol(s_export_selected)]

s_export_selected[12,2:ncol(s_export_selected)] <- s_export_selected[12,2:ncol(s_export_selected)] - s_export_selected[14,2:ncol(s_export_selected)]

s_export_selected <- s_export_selected %>%
  gather(key=year, value="export",-service)

ggplot(data = s_export_selected) +
  geom_histogram(aes(x = service))+
  coord_flip()
```

```{r}
s_export2 <- read_excel("data/SGTrade_byService2.xlsx",sheet = "T4",range="A11:k275")

s_export2 <- s_export2 %>%
  gather(key=year, value = "export", -c(Service,Country))%>%
  mutate(export = str_replace_all(export, "na", "0")) %>%
  mutate(export = as.numeric(export)) 
  
  
```

```{r}

s_export2 <- s_export2 %>%
  mutate(year = as.integer(year))

s_export2_2 <- s_export2 %>%
  filter(Country %in% c("United States Of America", "Hong Kong", "India", "Indonesia", "Japan", "Mainland China","Malaysia", "Philippines", "Republic Of Korea", "Thailand", "United Arab Emirates","France", "Germany", "Netherlands", "United Kingdom", "Australia"))



ggplot(s_export2_2, aes(x = export, 
                              y = Country,
                              fill = Country)) + 
  geom_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE,
                      quantiles = 4,
                      alpha = 0.6,
                      scale = 0.8) +
  geom_boxplot(outlier.shape = 16, 
               outlier.colour = "black",  
               outlier.size = 0.5,
               width = 0.2, 
               position = position_dodge(width = 0.7))+
  facet_grid(~Service,scales = "free_x")+
  theme_ridges() +
  labs(title = "",
       subtitle = "",
       x="Export Value" ,
       y="Country")+
  theme(
    axis.text.y = element_text(size = 7), 
    axis.text.x = element_text(size = 7),  
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    strip.text = element_text(size = 9, face = "bold"),  # Facet title size
    plot.margin = margin(10, 20, 10, 20),  
    legend.position = "none",
    legend.title = element_text(size=8),
    legend.key.size = unit(0.3, "cm"),
    legend.text = element_text(size=6),
    legend.background = element_rect(fill = "#f3f1e9"),
    panel.background = element_rect(fill = "#f3f1e9"),
    plot.title = element_text(size=13,hjust=0),
    plot.subtitle = element_text(size=8,hjust=0),
    plot.background = element_rect(fill = "#f3f1e9",color = NA)
    )
 # transition_time(year) + 
  #ease_aes('cubic-in-out')
```

```{r}
s_export2 <- s_export2 %>%
  mutate(year = as.integer(year))

# Create a treemap plot using ggplot2 (with geom_tile or geom_rect)
ggplot(s_export2, aes(x = Country, y = Service, size = export, fill = export)) +
  geom_tile() +  # or geom_rect() depending on your preference
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "RdYlBu")) +  # Use the palette
  scale_size_continuous(range = c(1, 20)) +  # Adjust the size range
  theme_void() +
  theme(legend.position = "right") +
  labs(title = "Export Data Treemap",
       subtitle = 'Year: {frame_time}',  # Subtitle dynamically changes with year
       fill = "Export Value") +
  transition_time(year) +  # Animate by year
  ease_aes('cubic-in-out')  # Smooth animation
```

::: panel-tabset
## Original

## Make-Over
:::

## 4 Reference

-   Kam, T.S. (2025).

<https://stackoverflow.com/questions/74015542/adding-a-shaded-triangle-to-a-ggplot2-plot>
