---
title: "Hands-on Exercise 08.c"
author: Johsuan Huang
date: 2024-03-13
date-modified: "last-modified"
description: "Analytical Mapping"
categories: [sf, tmap, tidyverse]
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

In this hands-on exercise, we are going to use appropriate functions of tmap and tidyverse to perform the following tasks:

-   Importing geospatial data in rds format into R environment.

-   Creating cartographic quality choropleth maps by using appropriate tmap functions.

-   Creating rate map

-   Creating percentile map

-   Creating boxmap

## 2 Getting Started

### 2.1 Load Packages

Use the **pacman** package `p_load()` to check, install and launch the following R packages:

```{r}
pacman::p_load(sf, tmap, tidyverse, knitr)
```

| **Library** | **Description** |
|------------------------------------|------------------------------------|
| **tidyverse** | a family of R packages for data processing |
| **sf** | For drawing thematic maps |
| **tmap** | Providing a standardized way to encode and analyze spatial vector data |

### 2.2 Import Data

For the purpose of this hands-on exercise, a prepared data set called *NGA_wp.rds* will be used. The data set is a polygon feature data.frame providing information on **water point of Nigeria at the LGA level**:

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

Let's check the data:

```{r}
#| echo: false
kable(head(NGA_wp))
```

## 3 Basic Choropleth Mapping

Plot the total water point and non-functional water point:

### 3.1 Static Map: Total WP vs Functional WP

```{r}
#| fig-width: 12


tmap_mode("plot")

p1 <- tm_shape(NGA_wp)+
  tm_fill("total_wp",
          n= 10,
          style = "equal",
          palette = "Purples") +
  tm_borders(lwd = 0.1,
             alpha = 0.2) +  
  tm_layout(main.title = "Distribution of total water point by LGAs",
            main.title.size = 2,
            main.title.position = "center",
            legend.position = c("right", "bottom"),
            legend.height = .4,
            legend.width = .2,
            legend.frame = FALSE,
            frame = FALSE,
            bg.color = "#f3f1e9")

p2 <- tm_shape(NGA_wp)+
  tm_fill("wp_functional",
          n=10,
          style="equal",
          palette ="Purples") +
  tm_borders(lwd = 0.1,
             alpha = 0.2) +  
  tm_layout(main.title = "Distribution of functional water point by LGAs",
            main.title.size = 2,
            main.title.position = "center",
            legend.position = c("right", "bottom"),
            legend.frame = FALSE,
            frame = FALSE,
            bg.color = "#f3f1e9")

tmap_arrange(p1, p2, nrow =1)
```

### 3.2 Interactive Map: Non-functional WP

```{r}

tmap_mode("view")

tm_shape(NGA_wp) +
  tm_basemap("Esri.WorldTopoMap") +
  tm_fill("wp_nonfunctional",
          n=10,
          style="equal",
          palette ="Purples") +
  tm_borders(lwd = 0.1, alpha = 0.3) +
  tm_shape(NGA_wp) +
  tm_polygons(alpha = 0,
              id = "ADM2_EN",
              popup.vars = c("State" = "ADM1_EN","Local Government Area" = "ADM2_EN", "Nonfunctional WP" = "wp_nonfunctional")) +
  tm_layout(main.title = "Distribution of nonfunctional water point by LGAs",
            main.title.size = 2,
            main.title.position = "center",
            legend.position = c("right", "bottom"),
            legend.frame = FALSE,
            frame = FALSE)
```

## 4 Choropleth Map for Rates

In much of our readings we have now seen the importance to map rates rather than counts of things, and that is for the simple reason that water points are not equally distributed in space. That means that if we do not account for how many water points are somewhere, we end up mapping total water point size rather than our topic of interest.

### 4.1 Deriving Proportion of Functional Water Points and Non-Functional Water Points

We will tabulate the proportion of functional water points and the proportion of non-functional water points in each LGA. In the following code chunk, `mutate(`) from **dplyr** package is used to derive two fields, namely *pct_functional* and *pct_nonfunctional*.

```{r}
NGA_wp_pct <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
```

### 4.2  Plotting map of rate

```{r}
tmap_mode("plot")

tm_shape(NGA_wp_pct) +
  tm_fill("pct_functional",
          n = 10,
          style = "equal",
          palette = "Blues",
          legend.hist = TRUE) +
  tm_borders(lwd = 0.1,
             alpha = .3) +
  tm_layout(main.title = "Rate map of functional water point by LGAs",
            main.title.size = 1,
            main.title.position = "center",          
            legend.outside = TRUE,
            frame = FALSE,
            legend.frame = FALSE,
            legend.text.size = 1,
            legend.hist.width = 0.7,
            bg.color = "#f3f1e9")
```

## 5 Extreme Value Maps

Extreme value maps are variations of common choropleth maps where the classification is designed to **highlight extreme values at the lower and upper end of the scale**, with the goal of identifying outliers. These maps were developed in the spirit of spatializing EDA, i.e., adding spatial features to commonly used approaches in non-spatial EDA (Anselin 1994).

### 5.1 Percentile Map

The percentile map is a special type of quantile map with six specific categories: 0-1%,1-10%, 10-50%,50-90%,90-99%, and 99-100%. The corresponding breakpoints can be derived by means of the base R quantile command, passing an explicit vector of **cumulative** **probabilities** as c(0,.01,.1,.5,.9,.99,1). Note that the begin and endpoint need to be included.

#### 5.1.1 Data Preparation

Step 1: Exclude records with NA by using the code chunk below.

```{r}
NGA_wp_dropna <- NGA_wp_pct %>%
  drop_na()
```

Step 2: Creating customised classification and extracting values

-   **st_set_geometry(NULL):** It removes the geometry from the `sf` object and leaves just the attribute data (columns in the data frame without the spatial information)

-   When variables are extracted from an sf data.frame, the geometry is extracted as well. For mapping and spatial manipulation, this is the expected behavior, but **many base R functions cannot deal with the geometry**. Specifically, the `quantile()` gives an error. As a result `st_set_geomtry(NULL)` is used to drop geomtry field.

```{r}
percent <- c(0,0.01,0.1,0.5,0.9,0.99,1)
var <- NGA_wp_dropna["pct_functional"] %>%
  st_set_geometry(NULL) #remove the geometry of a spatial object

quantile(var[,1], percent)
```

#### 5.1.2 Creating the get.var function

Firstly, we will write an R function as shown below to extract a variable (i.e. *wp_nonfunctional*) as a vector out of an sf data.frame.

-   arguments:

    -   vname: variable name (as character, in quotes)

    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

```{r}
get.var <- function(vname,df) {
  # Remove geometry
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  # Remove names from the column (unname)
  v <- unname(v[,1])
  
  return(v)
}
```

#### 5.1.3 A percentile mapping function

Next, we will write a percentile mapping function by using the code chunk below:

```{r}
tmap_mode("plot")

percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle,
            main.title.size = 2,
            main.title.position = c("center", "top"),
            legend.outside = TRUE,
            frame = FALSE,
            legend.frame = FALSE,
            legend.text.size = 1,
            legend.hist.width = 0.7,
            bg.color = "#f3f1e9")
}
```

\*\*Note: Since we have removed NA values, there should not be any missing values in the map

```{r}
#| fig-width: 10

percentmap("total_wp", NGA_wp_dropna)

```

### 5.2 Box map

In essence, a box map is an augmented quartile map, with an additional lower and upper category. When there are lower outliers, then the starting point for the breaks is the minimum value, and the second break is the lower fence. In contrast, when there are no lower outliers, then the starting point for the breaks will be the lower fence, and the second break is the minimum value (there will be no observations that fall in the interval between the lower fence and the minimum value).

-   Displaying summary statistics on a choropleth map by using the basic principles of boxplot.

-   To create a box map, a custom breaks specification will be used. However, there is a complication. The break points for the box map vary depending on whether lower or upper outliers are present.

#### 5.2.1 Creating the boxbreaks function

The code chunk below is an R function that creating break points for a box map.

-   arguments:

    -   v: vector with observations

    -   mult: multiplier for IQR (default 1.5)

-   returns:

    -   bb: vector with 7 break points compute quartile and fences

```{r}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr #Q3 + 1.5*IQR 
  lofence <- qv[2] - mult * iqr #Q1 - 1.5*IQR 
  
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

#### 5.2.2 Creating and Testing the get.var function

The code chunk below is an R function to extract a variable as a vector out of an sf data frame.

-   arguments:

    -   vname: variable name (as character, in quotes)

    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

Test drive the newly created function:

```{r}
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```

#### 5.2.3 Boxmap function

The code chunk below is an R function to create a box map. Below are arguments:

-   vnam: variable name (as character, in quotes)

-   df: simple features polygon layer

-   legtitle: legend title

-   mtitle: map title

-   mult: multiplier for IQR

-   returns: - a tmap-element (plots a map)

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(
            main.title = mtitle,
            main.title.size = 1.3,
            main.title.position = c("center", "top"),
            legend.outside = TRUE,
            frame = FALSE,
            legend.frame = FALSE,
            legend.text.size = 0.7,
            legend.hist.width = 0.7,
            bg.color = "#f3f1e9")
}
```

\*\*Note: Since we have removed NA values, there should not be any missing values in the map

```{r}
tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp_dropna)
```

## 6 Reference

-   Kam, T.S. (2025). [Analytical Mapping](https://r4va.netlify.app/chap23)
