---
title: "Hands-on Exercise 08.b"
author: Johsuan Huang
date: 2024-03-13
date-modified: "last-modified"
description: "Visualising Geospatial Point Data"
categories: [sf, tmap, tidyverse]
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

**Proportional symbol maps** (also known as graduate symbol maps) are a class of maps that use the visual variable of size to represent differences in the magnitude of a discrete, abruptly changing phenomenon, e.g. counts of people. Like choropleth maps, you can create classed or unclassed versions of these maps. The classed ones are known as range-graded or graduated symbols, and the unclassed are called proportional symbols, where the area of the symbols are proportional to the values of the attribute being mapped. In this hands-on exercise, you will learn how to create a proportional symbol map showing the number of wins by Singapore Pools’ outlets using an R package called **tmap**.

## 2 Getting Started

Use the **pacman** package `p_load()` to check, install and launch the following R packages:

```{r}
pacman::p_load(sf, tmap, tidyverse, knitr)
```

| **Library** | **Description** |
|----|----|
| **tidyverse** | a family of R packages for data processing |
| **sf** | For drawing thematic maps |
| **tmap** | Providing a standardized way to encode and analyze spatial vector data |

## 3 Geospatial Data Wrangling

The data set use for this hands-on exercise is called *SGPools_svy21*. The data is in csv file format.

It consists of seven columns. The XCOORD and YCOORD columns are the x-coordinates and y-coordinates of SingPools outlets and branches. They are in [Singapore SVY21 Projected Coordinates System](https://www.sla.gov.sg/sirent/CoordinateSystems.aspx).

### 3.1 Import data

```{r}
sgpools <- read_csv("data/aspatial/SGPools_svy21.csv")
```

After importing the data file into R, it is important for us to examine if the data file has been imported correctly:

```{r}
#| echo: false
kable(head(sgpools))
```

### 3.2 Creating a sf data frame from an aspatial data frame

The code chunk below converts sgpools data frame into a simple feature data frame by using *st_as_sf()* of **sf** packages

-   The ***coords*** argument requires you to provide the column name of the x-**coordinates** first then followed by the column name of the y-coordinates.

-   The ***crs*** argument required you to provide the **coordinates system** in epsg format. [EPSG: 3414](https://epsg.io/3414) is Singapore SVY21 Projected Coordinate System. You can search for other country’s epsg code by refering to [epsg.io](https://epsg.io/).

```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)
```

Table below shows the data table of *sgpools_sf*. Notice that a new column called **geometry(coordinates)** has been added into the data frame.

```{r}
kable(head(sgpools_sf))
```

## 4 Drawing Proportional Symbol Map

To create an interactive proportional symbol map in R, the view mode of tmap will be used.

### 4.1 Create an Interactive Point Symbol Map

Below are the arguments that can be edited in tm_bubbles():

-   **id:** shows information when hovering the mouse over a point

-   **popup.vars:** displays information when clicking on a point

The code chunks below are used to create an interactive point symbol map:

```{r}
tmap_mode("view")

tm_shape(sgpools_sf)+
  tm_basemap("Esri.WorldTopoMap") +
  tm_bubbles(col = "#E8C8DB",
           size = 1,
           border.col = "black",
           border.lwd = 1,
           id = "NAME",
           popup.vars=c("Name: " = "NAME",
                        "Type" = "OUTLET TYPE",
                        "Winnings: " = "Gp1Gp2 Winnings"))

```

### 4.2 Create an Interactive Proportional Point Symbol Map

To draw a proportional symbol map, we need to assign a numerical variable to the size visual attribute. The code chunks below show that the variable *Gp1Gp2Winnings* is assigned to size visual attribute:

```{r}
tm_shape(sgpools_sf)+
  tm_basemap("Esri.WorldTopoMap") +
  tm_bubbles(col = "#E8C8DB",
           size = "Gp1Gp2 Winnings",
           border.col = "black",
           border.lwd = 1,
           alpha = 0.9,
           id = "NAME",
           popup.vars=c("Name: " = "NAME",
                        "Type" = "OUTLET TYPE",
                        "Winnings: " = "Gp1Gp2 Winnings"))

```

### 4.3 Color by the attribute value

The proportional symbol map can be further improved by using the colour visual attribute. In the code chunks below, *OUTLET_TYPE* variable is used as the colour attribute variable:

```{r}
tm_shape(sgpools_sf)+
tm_basemap("Esri.WorldTopoMap") +
tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1,
          alpha = 0.9,
          id = "NAME",
          popup.vars=c("Name: " = "NAME",
                        "Address" = "ADDRESS",
                        "Winnings: " = "Gp1Gp2 Winnings"))
```

### 4.4Sync for faceted maps!

An impressive and little-know feature of **tmap**’s view mode is that it also works with faceted plots. The argument ***sync*** in *tm_facets()* can be used in this case to produce multiple maps with synchronised zoom and pan settings.

```{r}
tm_shape(sgpools_sf) +
  tm_basemap("Esri.WorldTopoMap") +
  tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1,
          id="NAME",
          popup.vars=c("Name: " = "NAME",
                        "Address" = "ADDRESS",
                        "Winnings: " = "Gp1Gp2 Winnings")) +
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

## 5 Reference

-   Kam, T.S. (2025). [Visualising Geospatial Point Data](https://r4va.netlify.app/chap22)
