---
title: "Hands-on Exercise 08.a"
author: Johsuan Huang
date: 2024-03-11
date-modified: "last-modified"
description: "Choropleth Mapping with R"
categories: [sf, tmap, tidyverse]
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors. For example, a social scientist may need to use a choropleth map to portray the spatial distribution of aged population of Singapore by Master Plan 2014 Subzone Boundary.

In this chapter, we will learn how to plot functional and truthful choropleth maps by using an R package called [**tmap**](https://cran.r-project.org/web/packages/tmap/) package.

## 2 Getting Started

### Load the packages

Use the **pacman** package `p_load()` to check, install and launch the following R packages:

```{r}
pacman::p_load(sf, tmap, tidyverse, knitr)
```

| **Library** | **Description** |
|----|----|
| **tidyverse** | a family of R packages for data processing |
| sf | For drawing thematic maps |
| tmap | Providing a standardized way to encode and analyze spatial vector data |

## 3 Import the data

Two data set will be used to create the choropleth map. They are:

-   Master Plan 2014 Subzone Boundary (Web) (i.e. `MP14_SUBZONE_WEB_PL`) in ESRI shapefile format. It can be downloaded at [data.gov.sg](https://data.gov.sg/) This is a geospatial data. It consists of the geographical boundary of Singapore at the planning subzone level. The data is based on URA Master Plan 2014.

-   Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format (i.e. `respopagesextod2011to2020.csv`). This is an aspatial data fie. It can be downloaded at [Department of Statistics, Singapore](https://www.singstat.gov.sg/) Although it does not contain any coordinates values, but it’s PA and SZ fields can be used as unique identifiers to geocode to `MP14_SUBZONE_WEB_PL` shapefile.

### 3.1 Import Geospatial data

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.

-   The `sf` package in R can read a wide variety of spatial file types. The `st_read()` function is built on GDAL (Geospatial Data Abstraction Library), allowing it to handle many vector data formats.

-   `dsn` stands for "data source name"

```{r}
mpsz <- st_read(dsn="data/geospatial",
                layer ="MP14_SUBZONE_WEB_PL")
```

```{r}
#| echo: false
kable(head(mpsz))
```

### 3.2 Import Aspatial Data

Next, we will import *respopagsex2011to2020.csv* file into RStudio and save the file into an R dataframe called *popagsex*.

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

```{r}
summary(popdata)
```

### 3.3 Data Preparation

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0_to_4 until age group 20_to_24

-   ECONOMY ACTIVE: age group 25_to_29 until age group 60_to_64

-   AGED: age group 65 and above

-   TOTAL: all age group

-   DEPENDENCY: the ratio between young and aged against economy active group

#### 3.3.1 Data wrangling

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package, and

-   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

    ```{r}
    popdata2020 <- popdata %>%
      filter(Time==2020) %>%
      group_by(PA,SZ,AG) %>%
      summarise(`POP`=sum(`Pop`)) %>%
      ungroup() %>%
      pivot_wider(names_from = AG,
                  values_from = POP) %>%
      mutate(YOUNG = rowSums(.[3:6])+rowSums(.[12]),
             ECONOMY_ACTIVE = rowSums(.[7:11])+rowSums(.[13:15]),
             AGED=rowSums(.[16:21]),
             TOTAL=rowSums(.[3:21]),
             DEPENDENCY = ((YOUNG+AGED)/ECONOMY_ACTIVE)) %>% 
               select(PA,SZ,YOUNG,AGED,ECONOMY_ACTIVE,TOTAL,DEPENDENCY)
      
    ```

#### 3.3.2 Join the aspatial data and geospatial data

Before we can perform the georelational join, one extra step is required to convert the values in PA and SZ fields to uppercase. This is because the values of PA and SZ fields are made up of upper- and lowercase. On the other, hand the SUBZONE_N and PLN_AREA_N are in uppercase.

```{r}
#| echo: true

popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>% #To uppercase
  filter(ECONOMY_ACTIVE > 0) #filter out NA data
```

Next, *left_join()* of **dplyr** is used to join the geographical data and attribute table using planning subzone name e.g. *SUBZONE_N* and *SZ* as the common identifier.

-   *left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
#| echo: false
kable(head(mpsz_pop2020))

```

#### 3.3.3 Export the geospatial data as rds file

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## 4 Choropleth Mapping Geospatial Data Using *tmap*

Two approaches can be used to prepare thematic map using *tmap*, they are:

-   Plotting a quick thematic map by using *qtm()*.

-   Plotting highly customisable thematic map by using tmap elements.

### 4.1 Plotting a choropleth map quickly by using *qtm()*

-   tmap_mode() has two modes:

    -   "**plot"**: This is the defalt mode, creating static thematic map

    -   "**view"**: This mode generate  a [`leaflet::leaflet()`](https://rstudio.github.io/leaflet/reference/leaflet.html) widget, creating interactive thematic map with map providers, such as OpenStreetMap.

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, fill="DEPENDENCY")

```

### 4.2 Creating a choropleth map by using *tmap*’s elements

Despite its usefulness of drawing a choropleth map quickly and easily, the disadvantge of *qtm()* is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, **tmap**’s drawing elements should be used.

```{r}
tmap_mode("plot")

tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Blues",
          title = "Dependency ratio")+
  tm_layout(main.title="Distribution of Dependency Ratio by planning subzone",
            main.title.position="center",
            main.title.size = 1.2,
            legend.frame = FALSE,
            legend.title.size = 0.7,
            legend.text.size = 0.6,
            bg.color = "#f3f1e9",
            frame = FALSE)+
  tm_borders(lwd = 0.5,alpha=0.5)+
  tm_compass(type="8star",size=2)+
  tm_scale_bar()+
  tm_grid(alpha = 0.2)+
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))

```

In the following sub-section, we will share with you tmap functions that used to plot these elements:

#### 4.2.1 Drawing a base map

The basic building block of **tmap** is *tm_shape()* followed by one or more layer elemments such as *tm_fill()* and *tm_polygons()*.

In the code chunk below, *tm_shape()* is used to define the input data (i.e *mpsz_pop2020*) and *tm_polygons()* is used to draw the planning subzone polygons:

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()+
  tm_layout(main.title="Base Map",
            main.title.position="center",
            bg.color = "#f3f1e9",
            frame = FALSE)
```

#### **4.2.2** Drawing a choropleth map using tm_polygons()

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")+
    tm_layout(main.title="Base Map with tm_polygons()",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

**Things to learn from *tm_polygons()*:**

-   The default interval binning used to draw the choropleth map is called “pretty”. A detailed discussion of the data classification methods supported by **tmap** will be provided in sub-section 4.3.

-   By default, Missing value will be shaded in grey.

#### 4.2.3 Drawing a choropleth map using tm_fill() and tm_border()

*tm_polygons()* is a wraper of *tm_fill()* and *tm_border()*. *tm_fill()* shades the polygons by using the default colour scheme and *tm_borders()* adds the borders of the shapefile onto the choropleth map.

-   *col* = border colour

-   *lwd* = border line width. The default is 1

-   *lty* = border line type. The default is “solid”.

-   *alpha* = adjsut border opacity (transparent) 0\~1 (opacity)

::: panel-tabset
## tm_fill()

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")+
    tm_layout(main.title="Base Map with tm_fill()",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## tm_border()

Notice that light-gray border lines have been added on the choropleth map:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="Base Map with tm_fill() & tm_border()",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```
:::

### 4.3 Data classification methods of tmap

Most choropleth maps employ some methods of **data classification**. The point of classification is to take a large number of observations and **group them into data ranges or classes**.

**tmap** provides a total ten data classification methods:

| classification method | Description |
|----|----|
| **fixed** | Manully define the breakpoints for classification. |
| **sd** | Classification is based on standard deviation from the mean. (highlight how much values deviate from mean) |
| **equal** | Divides the data range into equal-sized intervals. (best for evenly distributed data) |
| **pretty** | Ensures visually appealing class boundaries. |
| **quantile** | Divides data into bins where each class has **roughly the same number of observations**. |
| **kmeans** | Uses K-Means clustering to group values into `n` clusters. |
| **hclust** | Groups data into clusters based on hierarchical clustering. |
| **bclust** | Bagged Clustering: Similar to `hclust` but more robust by using bootstrap aggregating. |
| **fisher** | Finds natural breaks by minimizing variance within classes, deal for mapping geographical data where clusters naturally exist. |
| **jenks** | Good for highlighting natural groupings in data. |

#### 4.3.1 Plotting choropleth maps with built-in classification methods

To define a data classification method, the ***style*** argument of *tm_fill()* or *tm_polygons()* will be used:

::: panel-tabset
## janks

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5, #groups
          style = "jenks") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="jenks method: Good for highlighting natural groupings in data",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
  
```

## equal

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5, #groups
          style = "equal") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="equal method",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## quantile

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 4, #groups
          style = "quantile") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="quantile method",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## sd

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5, #groups
          style = "sd") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="sd method",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```
:::

#### 4.3.2 compare different n

Let use jenks method but with different numbers of classes (i.e. 2, 6, 10, 20) to see the difference across plots:

-   2 class: too generalized, showing only extreme values

-   6 class: comparisons become clearer

-   10 class: color difference are stronger than 6 classes and easier to interpret

-   20 class: too granular compared to 10 class, making the map less comparable

::: panel-tabset
## n=2

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 2, #groups
          style = "jenks") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="jenks method: 2 classes",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## n=6

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6, #groups
          style = "jenks") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="jenks method: 6 classes",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## n=10

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 10, #groups
          style = "jenks") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="jenks method: 10 classes",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## n=20

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 20, #groups
          style = "jenks") +
tm_borders(lwd = 0.5,  alpha = 0.5)+ #lwd = line width
    tm_layout(main.title="jenks method: 20 classes",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```
:::

#### 4.3.3 Plotting choropleth map with custome break

For all the built-in styles, the category breaks are computed internally. In order to override these defaults, the breakpoints can be set explicitly by means of the *breaks* argument to the *tm_fill()*. It is important to note that, **in tmap the breaks include a minimum and maximum**. As a result, in order to end up with n categories, n+1 elements must be specified in the *breaks* option (the values must be in increasing order).

Before we get started, it is always a good practice to get some descriptive statistics on the variable before setting the break points. Code chunk below will be used to compute and display the descriptive statistics of **DEPENDENCY** field:

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

With reference to the results above, we set break point at 0.60, 0.70, 0.80, and 0.90. In addition, we also need to include a minimum and maximum, which we set at 0 and 100. Our *breaks* vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00):

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)+
    tm_layout(main.title="self defined breaks: c(0, 0.60, 0.70, 0.80, 0.90, 1.00)",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## 4.4 Colour Scheme

**tmap** supports colour ramps either defined by the user or a set of predefined colour ramps from the **RColorBrewer** package.

### 4.4.1 ColourBrewer palette

To change the colour, we assign the preferred colour to *palette* argument of *tm_fill()* as shown in the code chunk below.

-   To reverse the colour shading, add a “-” prefix.

::: panel-tabset
## Greens

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Greens") +
  tm_borders(alpha = 0.4)+
  tm_layout(main.title="",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## Y10Br

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "YlOrBr") +
  tm_borders(alpha = 0.4)+
  tm_layout(main.title="",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```

## -virdus

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "-viridis") +
  tm_borders(alpha = 0.4)+
  tm_layout(main.title="",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame = FALSE)
```
:::

## 4.5 Map Layouts

Map layout refers to the combination of all map elements into a cohensive map. Map elements include among others the objects to be mapped, the title, the scale bar, the compass, margins and aspects ratios. Colour settings and data classification methods covered in the previous section relate to the palette and break-points are used to affect how the map looks.

### 4.5.1 Map Legend

In **tmap**, several *legend* options are provided to change the placement, format and appearance of the legend.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, # plot histogram for "dependency" only
          legend.is.portrait = TRUE, #forces the legend to be in a vertical (portrait) layout
          legend.hist.z = 0.1)+
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            frame = FALSE,
            legend.frame = FALSE,
            bg.color = "#f3f1e9") +
  tm_borders(lwd = 0.5,alpha = 0.3)
```

### 4.5.2 Map style

**tmap** allows a wide variety of layout settings to be changed. They can be called by using *tmap_style()*.

The code chunk below shows the *classic* style is used:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")+
    tm_layout(main.title="",
            main.title.position="center",
            legend.frame = FALSE,
            bg.color = "#f3f1e9")
```

### 4.5.3 Cartographic Furniture

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            frame = TRUE,
            legend.frame = FALSE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

To reset the default style, refer to the code chunk below.

```{r}
tmap_style("white")
```

## 4.6 Drawing Small Multiple Choropleth Maps

**Small multiple maps**, also referred to as **facet maps**, are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

In **tmap**, small multiple maps can be plotted in three ways:

-   by assigning multiple values to at least one of the asthetic arguments,

-   by defining a group-by variable in *tm_facets()*, and

-   by creating multiple stand-alone maps with *tmap_arrange()*.

### 4.6.1 By assigning multiple values to at least one of the aesthetic arguments

In this example, small multiple choropleth maps are created by defining ***ncols*** in **tm_fill()**

```{r}
#| fig-width: 12

map1 <- tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom"),
            legend.frame = FALSE,
            bg.color = "#f3f1e9"
            ) +
  tm_borders(alpha = 0.5)

tmap_arrange(map1, nrow = 1, ncol = 2)
```

In this example, small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments.

```{r}
#| fig-width: 12

map2 <- tm_shape(mpsz_pop2020) +
  tm_polygons("DEPENDENCY", 
              style = "equal", 
              palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom"),
            legend.frame = FALSE,
            bg.color = "#f3f1e9", frame=FALSE
            
            ) 

map3 <- tm_shape(mpsz_pop2020) +
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Greens") +
  tm_layout(legend.position = c("right", "bottom"),
            legend.frame = FALSE,
            bg.color = "#f3f1e9",
            frame=FALSE
            ) 

tmap_arrange(map2, map3, ncol = 1)

```

### 4.6.2 By defining a group-by variable in *tm_facets()*

In this example, multiple small choropleth maps are created by using **tm_facets()**

```{r}
mpsz_pop2020$REGION_N_CLEAN <- as.factor(trimws(as.character(mpsz_pop2020$REGION_N)))

tmap_mode("plot")

tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + #Keeps full polygon detail
  tm_facets(by="REGION_N",
            free.coords=TRUE,
            nrow = 3) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 4,
            frame = FALSE,
            bg.color = "#f3f1e9") +
  tm_borders(alpha = 0.5)
```

### 4.6.3 By creating multiple stand-alone maps with *tmap_arrange()*

-   asp: aspect ratio, if asp=1, meaning the maps will have equal width and height

```{r}
tmap_mode("plot")
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues",
              legend.frame = FALSE,
              bg.color = "#f3f1e9",
              frame=FALSE)

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues",
              legend.frame = FALSE,
              bg.color = "#f3f1e9",
              frame=FALSE)

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

Change to the interactive view mode:

```{r}
#| fig-width: 12

tmap_mode("view")
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues",
              legend.frame = FALSE,
              bg.color = "#f3f1e9",
              frame=FALSE,
              id = "SUBZONE_N",
              popup.vars = c("Subzone"="SUBZONE_N","Planning Area"="PLN_AREA_N","Young Pop"="YOUNG","Dependency %"="DEPENDENCY"))

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues",
              legend.frame = FALSE,
              bg.color = "#f3f1e9",
              frame=FALSE,
              id = "SUBZONE_N",
              popup.vars = c("Subzone"="SUBZONE_N","Planning Area"="PLN_AREA_N","Aged Pop"="AGED","Dependency %"="DEPENDENCY"))

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

## 4.7 Mappping Spatial Object Meeting a Selection Criterion

Instead of creating small multiple choropleth map, you can also use selection funtion to map spatial objects meeting the selection criterion.

```{r}
#| fig-width: 12


tmap_mode("plot")


tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = FALSE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.position = c("right", "bottom"),
            frame = FALSE,
            bg.color = "#f3f1e9",
            title = "Central Region") +
  tm_borders(alpha = 0.5)
```

# 5 Reference

-   Kam, T.S. (2025). [Choropleth Mapping with R](https://r4va.netlify.app/chap21)
