---
title: "Hands-on Exercise 05.d"
author: Johsuan Huang
date: 2024-02-12
date-modified: "last-modified"
description: "Treemap Visualisation with R"
categories: [treemap, treemapify, tidyverse]
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

In this hands-on exercise, you will gain hands-on experiences on designing treemap using appropriate R packages. The hands-on exercise consists of three main section. First, you will learn how to manipulate transaction data into a treemap strcuture by using selected functions provided in **dplyr** package. Then, you will learn how to plot static treemap by using **treemap** package. In the third section, you will learn how to design interactive treemap by using **d3treeR** package.

## 2 Getting Started

::: panel-tabset
## Loading the packages

Use the **pacman** package `p_load()` to check, install and launch the following R packages:

```{r}
pacman::p_load(treemap, treemapify, tidyverse)
```

| **Library** | **Description**                                              |
|------------------------|------------------------------------------------|
| tidyverse   | a family of R packages for data processing                   |
| treemapify  | The *treemapify* package allows creating treemaps in ggplot2 |

## Importing the data

In this hands-on exercise, *REALIS2018.csv* will be used. This dataset provides information of private property transaction records in 2018, and was extracted from [REALIS portal](https://spring.ura.gov.sg/lad/ore/login/index.cfm) of Urban Redevelopment Authority (URA).

```{r}
realis2018 <- read_csv("data/realis2018.csv")
```

## Observing the data

After importing the data, let's examine it to understand its data structure:

The dataset contains 20 attributes and 23,205 observations with no missing values.

```{r}
head(realis2018)
```

```{r}
str(realis2018)
```

```{r}
summary(realis2018)
```

```{r}
# check missing value
any(is.na(realis2018))
```

## Data Wrangling

The data.frame *realis2018* is in trasaction record form, which is highly disaggregated and not appropriate to be used to plot a treemap. In this section, we will perform the following steps to manipulate and prepare a data.frtame that is appropriate for treemap visualisation:

-   group transaction records by *Project Name*, *Planning Region*, *Planning Area*, *Property Type* and *Type of Sale*

-   compute *Total Unit Sold*, *Total Area*, *Median Unit Price* and *Median Transacted Price* by applying appropriate summary statistics on *No. of Units*, *Area (sqm)*, *Unit Price (\$ psm)* and *Transacted Price (\$)* respectively.

Two key verbs of **dplyr** package, namely: *group_by()* and *summarise()* will be used to perform these steps.

***group_by()*** breaks down a data.frame into specified groups of rows. When you then apply the verbs above on the resulting object they’ll be automatically applied “by group”.

Grouping affects the verbs as follows:

-   grouped *select()* is the same as ungrouped *select()*, except that grouping variables are always retained.

-   grouped *arrange()* is the same as ungrouped; unless you set *.by_group = TRUE*, in which case it orders first by the grouping variables.

-   *mutate()* and *filter()* are most useful in conjunction with window functions (like *rank()*, or *min(x) == x*).

-   *sample_n()* and *sample_frac()* sample the specified number/fraction of rows in each group.

-   *summarise()* computes the summary for each group.

In our case, *group_by()* will used together with *summarise()* to derive the summarised data.frame.

There are two ways to group the summaries:

#### Grouped summaries without the Pipe

The code chank below shows a typical two lines code approach to perform the steps.

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

#### Grouped summaries with the Pipe

The code chunk below shows a more efficient way to tackle the same processes by using the *pipe*, %\>%:

```{r}
realis2018_summarised <- realis2018 %>% 
  group_by(`Project Name`,`Planning Region`, 
           `Planning Area`, `Property Type`, 
           `Type of Sale`) %>%
  summarise(`Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE), 
            `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
            `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
            `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
head(realis2018_summarised)

```
:::

## 3 Designing Static Treemap with treemap Package

In this section, *`treemap()`* of **Treemap** package is used to plot a treemap showing the distribution of median unit prices and total unit sold of resale condominium by geographic hierarchy in 2017.

First, we will select records belongs to **resale & condominium** property type from *realis2018_selected* data frame using `filter()`.

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```

### 3.1 Using the basic arguments

The code chunk below designed a treemap by using three core arguments of *treemap()*, namely: *index*, *vSize* and *vColor*.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
  
```

::: {.callout-note appearance="simple"}
## Things to learn from the code chunk

-   **index**

    -   **The index vector must consist of at least two column names** or else no hierarchy treemap will be plotted.

    -   If multiple column names are provided, such as the code chunk above, the first name is the highest aggregation level, the second name the second highest aggregation level, and so on.(e.g. Central Region -\> Kallang -\> Pebble Bay)

-   **vSize**

    -   The column **must not contain negative values**. This is because its value will be used to map the **sizes of the rectangles** of the treemaps.
:::

### 3.2 Working with *vColor* and *type* arguments

For a correctly designed treemap, **the rectangles should display varying color intensities**. The plot in 3.1 is an incorrect example of this. In our case, **median unit prices** should determine the color intensity of each rectangle.

For *treemap()*, *vColor* is used in combination with the argument *type* to determines the colours of the rectangles. Without defining *type*, like the code chunk above, *treemap()* assumes type = index, in our case, the hierarchy of planning areas.

In the code chunk below, *type* argument is define as value.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type = "value",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

::: {.callout-note appearance="simple"}
## Thinking to learn from the conde chunk above

-   The rectangles are coloured with different intensity of green, reflecting their respective median unit prices.

-   The legend reveals that the values are binned into ten bins, i.e. 0-5000, 5000-10000, etc. with an equal interval of 5000.
:::

### 3.3 Colours in treemap package

There are two arguments that determine the mapping to color palettes: ***mapping*** and ***palette***.

The only difference between “value” and “manual” is the default value for mapping:

-   **"value" type treemap**: considers palette to be a diverging color palette (say ColorBrewer’s “RdYlBu”), and maps it in such a way that ***0*** **corresponds to the middle color** (typically white or yellow), *-max(abs(values))* to the left-end color, and *max(abs(values))*, to the right-end color.

-   **"manual" type treemap**: simply maps min(values) to the left-end color, max(values) to the right-end color, and mean(range(values)) to the middle color.

    :::: panel-tabset
    ## value-type

    The code chunk below shows a value type treemap:

    ```{r}
    treemap(realis2018_selected,
            index=c("Planning Region", "Planning Area", "Project Name"),
            vSize="Total Unit Sold",
            vColor="Median Unit Price ($ psm)",
            type="value",
            palette="RdYlBu", 
            title="Resale Condominium by Planning Region and Area, 2017",
            title.legend = "Median Unit Price (S$ per sq. m)"
            )
    ```

    ::: {.callout-note appearance="simple"}
    ## Thing to learn from the code chunk above

    -   although the colour palette used is RdYlBu but there are no red rectangles in the treemap above. This is because all the median unit prices are positive.

    -   The reason why we see only 5000 to 45000 in the legend is because the *`range`* argument is by default `c(min(values, max(values))` with some pretty rounding.
    :::

    ## manual-type

    The “manual” type does not interpret the values as the “value” type does. Instead, the value range is mapped linearly to the colour palette.

    As the plot shown below, The colour scheme used is very copnfusing. This is because mapping = (min(values), mean(range(values)), max(values)). **It is not wise to use diverging colour palette such as *RdYlBu* if the values are all positive or negative.**

    ```{r}
    treemap(realis2018_selected,
            index=c("Planning Region", "Planning Area", "Project Name"),
            vSize="Total Unit Sold",
            vColor="Median Unit Price ($ psm)",
            type="manual",
            palette="RdYlBu", 
            title="Resale Condominium by Planning Region and Area, 2017",
            title.legend = "Median Unit Price (S$ per sq. m)"
            )
    ```

    To overcome this problem, a single colour palette such as Blues should be used.

    ```{r}
    treemap(realis2018_selected,
            index=c("Planning Region", "Planning Area", "Project Name"),
            vSize="Total Unit Sold",
            vColor="Median Unit Price ($ psm)",
            type="manual",
            palette="Blues", 
            title="Resale Condominium by Planning Region and Area, 2017",
            title.legend = "Median Unit Price (S$ per sq. m)"
            )
    ```
    ::::

### 3.4 Treemap Layout

*treemap()* supports two popular treemap layouts, namely: “**squarified**” and “**pivotSize**”. The default is “pivotSize”.

The squarified treemap algorithm (Bruls et al., 2000) produces good aspect ratios, but ignores the sorting order of the rectangles (sortID). The ordered treemap, pivot-by-size, algorithm (Bederson et al., 2002) takes the sorting order (sortID) into account while aspect ratios are still acceptable.

### 3.5 Working with *algorithm* argument

The code chunk below plots a squarified treemap by changing the *algorithm*argument:

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "squarified",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

### 3.6 Using *sortID*

When “pivotSize” algorithm is used, *sortID* argument can be used to dertemine the order in which the rectangles are placed from top left to bottom right.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## 4 Designing Treemap using treemapify Package

**treemapify** is a R package specially developed to draw treemaps in **ggplot2**. In this section, you will learn how to designing treemps closely resemble treemaps designing in previous section by using **treemapify**. Before you getting started, you should read [Introduction to “treemapify”](https://cran.r-project.org/web/packages/treemapify/vignettes/introduction-to-treemapify.html) its [user guide](https://cran.r-project.org/web/packages/treemapify/treemapify.pdf).

### 4.1 Designing a basic treemap

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`),
       layout = "scol",
       start = "bottomleft") + 
  geom_treemap() +
  scale_fill_gradient(low = "light blue", high = "blue")
```

### 4.2 Defining hierarchy

::: panel-tabset
## Group by Planning Region

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`),
       start = "topleft") + 
  geom_treemap()
```

## Group by Planning Area

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) +  #<<< adding
  geom_treemap()
```

Adding boundary line:

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40", #<< adding border for PA
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20") #<< adding border for planning region
```
:::

## 5 Designing Interactive Treemap using d3treeR

### 5.1 Load *d3treeR* package

```{r}
library(d3treeR)

```

### 5.2 Designing An Interactive Treemap

The codes below perform two processes.

1.  *treemap()* is used to build a treemap by using selected variables in condominium data.frame. The treemap created is save as object called ***tm***.

    ```{r}
    tm <- treemap(realis2018_summarised,
            index=c("Planning Region", "Planning Area"),
            vSize="Total Unit Sold",
            vColor="Median Unit Price ($ psm)",
            type="value",
            title="Private Residential Property Sold, 2017",
            title.legend = "Median Unit Price (S$ per sq. m)"
            )
    ```

<!-- -->

2.  Then *`d3tree()`* is used to build an interactive treemap:

    ```{r}
    d3tree(tm,rootname = "Singapore") #<< set up root name (the root) as Singapore
    ```

## 6 Reference

-   Kam, T.S. (2025). [Treemap Visualisation with R](https://r4va.netlify.app/chap16 "Treemap Visualisation with R")
