---
title: "Hands-on Exercise 09"
author: Johsuan Huang
date: 2024-03-22
date-modified: "last-modified"
description: "Modelling, Visualising and Analysing Network Data with R"
categories: [igraph, tidygraph, ggraph, visNetwork, lubridate, clock,
               tidyverse, graphlayouts]
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

In this hands-on exercise, we will learn how to model, analyse and visualise network data using R:

-   create graph object data frames, manipulate them using appropriate functions of *dplyr*, *lubridate*, and *tidygraph*,

-   build network graph visualisation using appropriate functions of *ggraph*,

-   compute network geometrics using *tidygraph*,

-   build advanced graph visualisation by incorporating the network geometrics, and

-   build interactive network visualisation using *visNetwork* package.

# 2 Getting Started

## Load the packages

In this hands-on exercise, four network data modelling and visualisation packages will be installed and launched. They are igraph, tidygraph, ggraph and visNetwork. Beside these four packages, tidyverse and [lubridate](https://lubridate.tidyverse.org/), an R package specially designed to handle and wrangling time data will be installed and launched too.

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts,knitr)
```

# 3 Import the data

The data sets used in this hands-on exercise is from an oil exploration and extraction company. There are two data sets One contains the nodes data and the other contains the edges (also know as link) data.

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")
```

#### The edges data

-   *GAStech-email_edges.csv* which consists of two weeks of 9063 emails correspondances between 55 employees.

```{r}
#| echo: false
kable(head(GAStech_edges))
```

#### The nodes data

-   *GAStech_email_nodes.csv* which consist of the names, department and title of the 55 employees.

```{r}
kable(head(GAStech_nodes))
```

## 3.1 Check the data structure

The output report of GAStech_edges above reveals that the *SentDate* is treated as “Character” data type instead of *date* data type. This is an error! Before we continue, it is important for us to change the data type of *SentDate* field back to “Date” data type.

```{r}
str(GAStech_edges)
```

```{r}
str(GAStech_nodes)
```

## 3.2 Wrangling Time

The code chunk below will be used to perform the changes:

-   *wday()* returns the day of the week as a decimal number by default

    -   label=TRUE: returns the character version of the ordered labels instead of numbers

    -   abbr=TRUE: returns the abbreviated label (e.g., MON, TUE)

-   the values in the *Weekday* field are in ordinal scale.

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate)) %>% #Day,Month,Year
  mutate(Weekday = wday(SentDate, 
                        label = TRUE, # return Chr version of data
                        abbr = FALSE)) # No Abbreviation
```

Let's check the revised data:

```{r}
kable(head(GAStech_edges))
```

## 3.3 Wrangling Attributes

A close examination of *GAStech_edges* data.frame reveals that it consists of individual e-mail flow records. This is not very useful for visualisation.

In view of this, we will aggregate the individual by date, senders, receivers, main subject and day of the week.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n(),.groups = "drop") %>% # count occurrences as Weight
  filter(source!=target) %>% # remove emails where the sender and receiver are the same person
  filter(Weight > 1) %>%
  ungroup()
```

Below table shows the reformed *GAStech_edges* data frame

```{r}
kable(head(GAStech_edges_aggregated))
```

# 4 Creating network objects using tidygraph

**tidygraph** package provides a tidy API for graph/network manipulation. While network data itself is not tidy, it can be envisioned as two tidy tables, one for **node data** and one for **edge data**.

tidygraph provides a way to switch between the two tables and provides dplyr verbs for manipulating them. Furthermore it provides access to a lot of graph algorithms with return values that facilitate their use in a tidy workflow.

## 4.1 The tbl_graph object

Two functions of **tidygraph** package can be used to create network objects, they are:

-   [`tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html) creates a **tbl_graph** network object from nodes and edges data.

-   [`as_tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html) converts network data and objects to a **tbl_graph** network. Below are network data and objects supported by `as_tbl_graph()`

    -   a node data.frame and an edge data.frame,

    -   data.frame, list, matrix from base,

    -   igraph from igraph,

    -   network from network,

    -   dendrogram and hclust from stats,

    -   Node from data.tree,

    -   phylo and evonet from ape, and

    -   graphNEL, graphAM, graphBAM from graph (in Bioconductor).

## 4.2 The dplyr verbs in tidygraph

-   *activate()* verb from **tidygraph** serves as a switch between tibbles for nodes and edges. All dplyr verbs applied to **tbl_graph** object are applied to the active tibble.

    ``` r
    iris_tree <- iris_tree %>% 
      activate(nodes) %>% 
      # Create new column Species if a node is a leaf, and assign label to it
      mutate(Species = ifelse(leaf, as.character(iris$Species)[label], NA)) %>% 
      activate(edges) %>% 
      # .N()$Species[to] refer to the node the edge is pointing to >> setosa, which will return TRUE/FALSE
      mutate(to_setosa = .N()$Species[to] =='setosa')
    ```

-   In the above the *.N()* function is used to gain access to the node data while manipulating the edge data. Similarly *.E()* will give you the edge data and *.G()*will give you the **tbl_graph** object itself.

## 4.3 Using `tbl_graph()` to build tidygraph data model

-   **`directed = TRUE`**: Specifies that the graph is **directed** (edges have a specific direction from `source` to `target`).

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
```

Reviewing GAStech_graph table:

-   The output below reveals that *GAStech_graph* is a tbl_graph object with 54 nodes and 4541 edges, **meaning there are 54 unique persons with 4541 directed connections between nodes**.

-   The command also prints the first six rows of “Node Data” and the first three of “Edge Data”.

-   It states that the Node Data is **active (Node data is on top of Edge data)**. The notion of an active tibble within a `tbl_graph` object makes it possible to manipulate the data in one tibble at a time.

```{r}
GAStech_graph
```

## 4.4 Changing the sort order of the active object

The nodes tibble data frame is activated by default, but you can change which tibble data frame is active with the *activate()* function. Thus, if we wanted to rearrange the rows in the edges tibble to list those with the highest “weight” first, we could use *activate()* and then *arrange()*.

```{r}
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight)) # arrange the order of edges by desc weight

```

# 5 Plotting Static Network Graphs with ggraph package

[**ggraph**](https://ggraph.data-imaginist.com/) is an extension of **ggplot2**, making it easier to carry over basic ggplot skills to the design of network graphs.

As in all network graph, there are three main aspects to a **ggraph**’s network graph, they are:

-   [nodes](https://cran.r-project.org/web/packages/ggraph/vignettes/Nodes.html),

-   [edges](https://cran.r-project.org/web/packages/ggraph/vignettes/Edges.html) and

-   [layouts](https://cran.r-project.org/web/packages/ggraph/vignettes/Layouts.html).

## 5.1 Plotting a basic network graph

The code chunk below uses [*ggraph()*](https://ggraph.data-imaginist.com/reference/ggraph.html), [*geom-edge_link()*](https://ggraph.data-imaginist.com/reference/geom_edge_link.html) and [*geom_node_point()*](https://ggraph.data-imaginist.com/reference/geom_node_point.html) to plot a network graph by using *GAStech_graph*:

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

::: {.callout-note appearance="simple"}
-   The basic plotting function is `ggraph()`, which takes the data to be used for the graph and the type of layout desired. Both of the arguments for `ggraph()` are built around *igraph*. Therefore, `ggraph()` can use either an *igraph* object or a *tbl_graph* object.
:::

## 5.2 Changing the default network graph theme

In this section, you will use [*theme_graph()*](https://ggraph.data-imaginist.com/reference/theme_graph.html) to remove the x and y axes. Before your get started, it is advisable to read it’s reference guide at least once.

```{r}
#| fig-height: 6

g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes(colour = factor(Weekday))) +
  geom_node_point(aes())+
  geom_node_text(color = "black",
                 fontface= "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))

g + theme_graph()+
  labs(title="Email Network Analysis by Day of Week")+
  theme(
      text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
      plot.title = element_text(hjust = 0.5),
      legend.position = 'top',
      legend.background = element_rect(fill = "#f3f1e9", color = NA),
      legend.key.size = unit(0.2, "cm"),
      legend.title = element_text(size = 9),
      legend.text = element_text(size = 8),
      panel.background = element_rect(fill = "#f3f1e9", color = NA),
      plot.background = element_rect(fill = "#f3f1e9", color = NA))
```

::: {.callout-note appearance="simple"}
-   **ggraph** introduces a special ggplot theme that provides better defaults for network graphs than the normal ggplot defaults. `theme_graph()`, besides removing axes, grids, and border, changes the font to Arial Narrow (this can be overridden).

-   The ggraph theme can be set for a series of plots with the `set_graph_style()`command run before the graphs are plotted or by using `theme_graph()` in the individual plots.
:::

## 5.3 Changing the coloring of the plot

```{r}

g <- ggraph(GAStech_graph) + 
  geom_edge_link(colour = "grey80") +   # Edge color
  geom_node_point(colour="#8087CA") +   # Node color
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels

g + 
  theme_graph() +
  labs(title = "Email Network Analysis") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5),  
    legend.position = "none", 
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA)  
  )

```

## 5.4 Working with different ggraph’s layouts

**ggraph** support many layout for standard used, they are: star, circle, nicely (default), dh, gem, graphopt, grid, mds, spahere, randomly, fr, kk, drl and lgl. Figures below and on the right show layouts supported by `ggraph()`.

::: panel-tabset
## FR

Fruchterman and Reingold layout:

```{r}

g <- ggraph(GAStech_graph,layout = "fr") + 
  geom_edge_link(colour = "grey80") +   # Edge color
  geom_node_point(colour="#8087CA") +   # Node color
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels

g + 
  theme_graph() +
  labs(title = "Email Network Analysis") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "none",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

## star

```{r}
g <- ggraph(GAStech_graph,layout = "star") + 
  geom_edge_link(colour = "grey80") +   # Edge color
  geom_node_point(colour="#8087CA") +   # Node color
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels

g + 
  theme_graph() +
  labs(title = "Email Network Analysis") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "none",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
  
```

## sphere

```{r}
g <- ggraph(GAStech_graph,layout = "sphere") + 
  geom_edge_link(colour = "grey80") +   # Edge color
  geom_node_point(colour="#8087CA") +   # Node color
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels

g + 
  theme_graph() +
  labs(title = "Email Network Analysis") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "none",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

## graphopt

```{r}
g <- ggraph(GAStech_graph,layout = "graphopt") + 
  geom_edge_link(colour = "grey80") +   # Edge color
  geom_node_point(colour="#8087CA") +   # Node color
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels

g + 
  theme_graph() +
  labs(title = "Email Network Analysis") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "none",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

## grid

```{r}
g <- ggraph(GAStech_graph,layout = "grid") + 
  geom_edge_link(colour = "grey80") +   # Edge color
  geom_node_point(colour="#8087CA") +   # Node color
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels

g + 
  theme_graph() +
  labs(title = "Email Network Analysis") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "none",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```
:::

## 5.5 Modifying network nodes

In this section, you will colour each node by referring to their respective departments.

-   *geom_node_point* is equivalent in functionality to *geo_point* of **ggplot2**. It allows for simple plotting of nodes in different shapes, colours and sizes. In the codes chnuks above colour and size are used.

```{r}
#| fig-height: 6

g <- ggraph(GAStech_graph,layout = "nicely") + 
  geom_edge_link(colour = "black",
                 size = 0.1) +  
  geom_node_point(aes(colour = Department),
                  size = 3,
                  shape = 18) 

g + 
  theme_graph() +
  labs(title = "Email Network Analysis by Department") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "top",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

## 5.6 Modifying Edges

In the code chunk below, the thickness of the edges will be mapped with the *Weight* variable.

-   *geom_edge_link* draws edges in the simplest way - as straight lines between the start and end nodes. But, it can do more that that. In the example above, argument *width* is used to map the width of the line in proportional to the Weight attribute and argument alpha is used to introduce opacity on the line.

```{r}
#| fig-height: 6

g <- ggraph(GAStech_graph,layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +  
  geom_node_point(aes(colour = Department),
                  size = 3,
                  shape = 18) 

g + 
  theme_graph() +
  labs(title = "Email Network Analysis by Department") +
  theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "top",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

# 6 Creating facet graphs

Another very useful feature of **ggraph** is faceting. In visualising network data, this technique can be used to reduce edge over-plotting in a very meaning way by spreading nodes and edges out based on their attributes. In this section, you will learn how to use faceting technique to visualise network data.

There are three functions in ggraph to implement faceting, they are:

-   [*facet_nodes()*](https://r4va.netlify.app/chap27) whereby edges are only draw in a panel if both terminal nodes are present here,

-   [*facet_edges()*](https://ggraph.data-imaginist.com/reference/facet_edges.html) whereby nodes are always drawn in al panels even if the node data contains an attribute named the same as the one used for the edge facetting, and

-   [*facet_graph()*](https://ggraph.data-imaginist.com/reference/facet_graph.html) faceting on two variables simultaneously.

## 6.1 Working with *facet_edges()*

In the code chunk below, [*facet_edges()*](https://ggraph.data-imaginist.com/reference/facet_edges.html) is used. Before getting started, it is advisable for you to read it’s reference guide at least once.

```{r}
#| fig-height: 6

set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.5) +
  scale_edge_width(range = c(0.1, 5)) +   # edge widths range from 0.1 to 5 
  geom_node_point(aes(colour = Department), 
                  size = 2)

g + facet_edges(~Weekday)+
    labs(title = "Email Network Analysis by Weekday") +
    theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "top",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
  
```

## 6.2 A framed facet graph

```{r}
#| fig-height: 6

set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
    labs(title = "Email Network Analysis by Weekday") +
    theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "top",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

## 6.3 Working with *facet_nodes()*

In the code chunkc below, [*facet_nodes()*](https://ggraph.data-imaginist.com/reference/facet_nodes.html) is used. Before getting started, it is advisable for you to read it’s reference guide at least once.

```{r}
#| fig-height: 6

set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_nodes(~Department) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
    labs(title = "Email Network Analysis by Department") +
    theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "top",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

# 7 Network Metrics Analysis

## 7.1 Computing centrality indices

[Centrality measures are a collection of statistical indices use to describe the relative important of the actors are to a network]{.underline}. There are four well-known centrality measures, namely: degree, betweenness, closeness and eigenvector. It is beyond the scope of this hands-on exercise to cover the principles and mathematics of these measure here. (Students are encouraged to refer to ***Chapter 7: Actor Prominence*** of **A User’s Guide to Network Analysis in R** to gain better understanding of theses network measures)

-   *mutate()* of **dplyr** is used to perform the computation.

-   the algorithm used, on the other hand, is the *centrality_betweenness()* of **tidygraph**.

```{r}
#| fig-height: 6

g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2,
                 colour = "grey30") +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
            size=betweenness_centrality))+
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels
g + theme_graph()+
    labs(title = "Email Network Analysis by Department and Centrality") +
    theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "right",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

## 7.2 Visualising network metrics

It is important to note that from **ggraph v2.0** onward tidygraph algorithms such as centrality measures can be accessed directly in ggraph calls. This means that it is no longer necessary to precompute and store derived node and edge centrality measures on the graph in order to use them in a plot.

```{r}
#| fig-height: 6

g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2,
                 colour = "grey30") +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
            size=centrality_betweenness()))+ # Buit-in 
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels
g + theme_graph()+
    labs(title = "Email Network Analysis by Department and Centrality") +
    theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "right",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

## 7.3 Visualising Community

tidygraph package inherits many of the community detection algorithms imbedded into igraph and makes them available to us, including *Edge-betweenness (group_edge_betweenness)*, *Leading eigenvector (group_leading_eigen)*, *Fast-greedy (group_fast_greedy)*, *Louvain (group_louvain)*, *Walktrap (group_walktrap)*, *Label propagation (group_label_prop)*, *InfoMAP (group_infomap)*, *Spinglass (group_spinglass)*, and *Optimal (group_optimal)*.

Some community algorithms are designed to take into account direction or weight, while others ignore it. Use this [link](https://tidygraph.data-imaginist.com/reference/group_graph.html) to find out more about community detection functions provided by tidygraph,

In the code chunk below *group_edge_betweenness()* is used:

```{r}
#| fig-height: 6

g <- GAStech_graph %>%
  mutate(community = as.factor(group_edge_betweenness(weights = Weight, 
                                                      directed = TRUE))) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2,
                 colour = "grey30") +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = community))+
  geom_node_text(colour = "black",
                 fontface = "bold",
                 size = 2,
                 repel = TRUE,
                 aes(label = label))  # Node labels
g + theme_graph()+
    labs(title = "Email Network Analysis by Community") +
    theme(
    text = element_text(family = "Arial", face = "plain", size = 11, color = "black"),
    plot.title = element_text(hjust = 0.5), 
    legend.position = "right",  
    legend.background = element_rect(fill = "#f3f1e9", color = NA),
    legend.key.size = unit(0.2, "cm"),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.background = element_rect(fill = "#f3f1e9", color = NA),
    plot.background = element_rect(fill = "#f3f1e9", color = NA) 
  )
```

# 8 Building Interactive Network Graph with visNetwork

-   [visNetwork()](http://datastorm-open.github.io/visNetwork/) is a R package for network visualization, using [vis.js](http://visjs.org/) javascript library.

-   *visNetwork()* function uses a nodes list and edges list to create an interactive graph.

    -   The nodes list must include an “id” column, and the edge list must have “from” and “to” columns.

    -   The function also plots the labels for the nodes, using the names of the actors from the “label” column in the node list.

-   The resulting graph is fun to play around with.

    -   You can move the nodes and the graph will use an algorithm to keep the nodes properly spaced.

    -   You can also zoom in and out on the plot and move it around to re-center it.

## 8.1 Data preparation

Before we can plot the interactive network graph, we need to prepare the data model by using the code chunk below.

```{r}
# aggregated data by from and to
GAStech_edges_aggregated_2 <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel" = "label")) %>%
  rename(from = id) %>%
  left_join(GAStech_nodes, by = c("targetLabel" = "label")) %>%
  rename(to = id) %>%
  filter(MainSubject == "Work related") %>%
  group_by(from, to) %>%
    summarise(weight = n(),.groups = "drop") %>%
  filter(from!=to) %>%
  filter(weight > 1) %>%
  ungroup()
```

## 8.2 Plotting the interactive network graph

The code chunk below will be used to plot an interactive network graph by using the data prepared.

```{r}
visNetwork(GAStech_nodes, 
           GAStech_edges_aggregated_2,
           background = "#f3f1e9")
```

```{r}
visNetwork(GAStech_nodes, 
           GAStech_edges_aggregated_2,
           background = "#f3f1e9")%>%
  visIgraphLayout(layout = "layout_with_fr") 

```

Visit [Igraph](http://datastorm-open.github.io/visNetwork/igraph.html) to find out more about *visIgraphLayout*’s argument.

## 8.3 Working with visual attributes - Nodes

visNetwork() looks for a field called “group” in the nodes object and colour the nodes according to the values of the group field.

The code chunk below rename Department field to group.

```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group = Department) 
```

When we return the code chunk below, visNetwork shades the nodes by assigning unique colour to each category in the *group* field.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated_2,
           background = "#f3f1e9") %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

## 8.4 Working with visual attributes - Edges

In the code run below *visEdges()* is used to symbolise the edges.\
- The argument *arrows* is used to define where to place the arrow.\
- The *smooth* argument is used to plot the edges using a smooth curve.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated_2,
           background = "#f3f1e9") %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visEdges(arrows = "to",                        # arrow to "to"
           smooth = list(enabled = TRUE, 
                         type = "curvedCW")) %>% # Smooth curve
  visLegend() %>%
  visLayout(randomSeed = 123)
```

Visit [Option](http://datastorm-open.github.io/visNetwork/edges.html) to find out more about visEdges’s argument.

## 8.6 Interactivity

In the code chunk below, *visOptions()* is used to incorporate interactivity features in the data visualisation.

-   The argument *highlightNearest* highlights nearest when clicking a node.

-   The argument *nodesIdSelection* adds an id node selection creating an HTML select element.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated_2,
           background = "#f3f1e9") %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

# 9 Reference

-   Kam, T.S. (2025). [Modelling, Visualising and Analysing Network Data with R](https://r4va.netlify.app/chap27)
