---
title: "Take-Home Exercise 3: Prototyping Modules for Visual Analytics Shiny Application"
author: Johsuan Huang
date: 2024-03-21
date-modified: "last-modified"
categories: []
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 The Task

In this take-home exercise, you are required to select one of the module of your proposed Shiny application and complete the following tasks:

-   To evaluate and determine the necessary R packages needed for your Shiny application are supported in R CRAN,
-   To prepare and test the specific R codes can be run and returned the correct output as expected,
-   To determine the parameters and outputs that will be exposed on the Shiny applications, and
-   To select the appropriate Shiny UI components for exposing the parameters determine above.

# 2 Getting Started

Our group project focuses on Singapore's weather, and we aim to provide an interactive weather dashboard where users can explore and gain a deeper understanding of Singapore's weather trends.

I will be working on two modules: **Co-variance Analysis** and **Forecasting**.

For a better understanding of these modules, please refer to our project proposal: [**Wheather or Not – Predicting the Unpredictable**](https://vaagroup13.netlify.app/team/project_proposal)

## 2.1 Load Packages

Below are the packages I'll be using for testing and building the prototypes:

```{r}
pacman::p_load(DT,sf, tmap, tidyverse,urca,tseries,plotly,ggdist,ggridges,lubridate,knitr,zoo)
```

| Package | Description |
|------------------------------------|------------------------------------|
| [tidyverse](https://www.tidyverse.org/) | For data manipulation |
| [DT](https://rstudio.github.io/DT/) | For interactive data tables |
| [knitr](https://yihui.org/knitr/) | For dynamic report generation with R |
| [ggstatplot](https://indrajeetpatil.github.io/ggstatsplot/) | To visual correlation matrix and plots with statistical tests |
| [plotly](https://plotly.com/r/) | For interactive, publication-quality plots |
| [ggiraph](https://davidgohel.github.io/ggiraph/) | For dynamic, interactive ggplot2 graphs. |
| [scales](https://scales.r-lib.org/) | For data labels and annotations for ggplot2 |
| [patchwork](https://patchwork.data-imaginist.com/) | For preparing composite figure created using ggplot2. |
| [ggdist](https://mjskay.github.io/ggdist/) | For visualisations of distributions and uncertainty. |
| [ggtext](https://wilkelab.org/ggtext/) | Providing simple Markdown and HTML rendering for ggplot2 |
| [ggridges](https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html) | For ridgeline plots that create the impression of a mountain range. They can be useful for visualising changes in distributions over time or space. |
| [lubridate](https://lubridate.tidyverse.org/) | For handling date-time data |
| [tsibble](https://tsibble.tidyverts.org/) | For tidy temporal data with wrangling tools |
| [feasts](https://feasts.tidyverts.org/) | For analysing tidy time series data with decomposition methods (STL), statistical summaries and graphics functions(cycle plot) |
| [fable](https://fable.tidyverts.org/) | Forecasting models for tidy time series data |
| [tmap](https://r-tmap.github.io/tmap/reference/index.html) | For drawing thematic maps |
| [sf](https://r-spatial.github.io/sf/) | Providing a standardized way to encode and analyze spatial vector data |
| [urca](https://www.pfaffikus.de/rpacks/urca/) | For co-integration analysis (Engle-Granger and Johansen tests) |

## 2.2 Import and Pre-process Weather Data

```{r}
station<-read_csv("data/Singapore_daily_records.csv")
```

### 2.2.1 Review the data structure

There are 22,230 observation

As shown in the table below, several variables have incorrect data formats, the transformation needed to be made for these variables:

-   station should be a factor type

-   rainfall, temperature, and wind speed should be numeric types

-   highest 30, 60 and 90 mm rainfall columns contain character "\\u0997", which should be replaced by NA

```{r}
str(station)
```

### 2.2.2 Handle Incorrect Data Format

The code chunk below is performed to address the data issues

```{r}
station <- station %>%
  mutate(across(
    c(highest_30_min_rainfall_mm, highest_60_min_rainfall_mm, highest_120_min_rainfall_mm,
      daily_rainfall_total_mm, mean_temperature_c, maximum_temperature_c, 
      minimum_temperature_c, mean_wind_speed_km_h, max_wind_speed_km_h,mean_temperature_a_c,
      maximum_temperature_a_c,minimum_temperature_a_c),
    ~ as.numeric(gsub("\u0097", NA, .))
  )) %>%
  mutate(station = as.factor(station))
```

### 2.2.3 Add Date and Week No.

As the weather data is a time-series data, it's necessary to add date information for further data analysis. The `lubridate` package is used for create date and weekday columns:

```{r}
station <- station %>%
  mutate(date = ymd(paste(year, month, day, sep = "-")),
         weekday = wday(date, label = TRUE, abbr = TRUE), 
         week_no = week(date))

# if week_no =53, revise it to 52
station <- station %>%
  mutate(week_no = ifelse(week_no == 53, 52, week_no))
```

### 2.2.4 Handle Missing Values

Now, we can check the data summary and missing value:

Except for columns "station", "year", "month" and "day", the data contain lots of missing values, especially for temperature data (\*\_temperature_c and \*\_temperature_ac)

```{r}
summary(station)
```

```{r}
kable(data.frame(Column = names(station), NAs = colSums(is.na(station))))

```

Upon closer inspection of the missing values, we discovered that the columns were switched: missing temperature_c values appeared in temperature_a_c, and vice versa. To maintain data integrity, we will use the `coalesce()` function to combine these values into the temperature_c column.

::: panel-tabset
## missing temperature_c

The table below shows records with missing values in mean_temperature_c

```{r}
station_filtered_c <- station %>%
  filter(is.na(mean_temperature_c))

kable(head(station_filtered_c,10))
```

## missing temperature_a_c

The table below shows records with missing values in mean_temperature_a_c

```{r}
station_filtered_a_c <- station %>%
  filter(is.na(mean_temperature_a_c))

kable(head(station_filtered_a_c,10))
```
:::

The coalesce function combines values from min, max, and mean temperature columns (both temp_c and temp_a_c) to handle missing values, and remove the \_a_c columns:

```{r}
station$mean_temperature_c <- coalesce(station$mean_temperature_c, station$mean_temperature_a_c)
station$maximum_temperature_c <- coalesce(station$maximum_temperature_c, station$maximum_temperature_a_c)
station$minimum_temperature_c <- coalesce(station$minimum_temperature_c, station$minimum_temperature_a_c)

station <- station %>% 
  select(-c("mean_temperature_a_c", 
            "maximum_temperature_a_c", "minimum_temperature_a_c"))

```

Now, let's review the summary of missing values again. The number of missing values has dramatically decreased to 1000ish.

```{r}
kable(data.frame(Column = names(station), NAs = colSums(is.na(station))))
```

As shown in the time-series plot, Marina Barrage has missing data since later half of year in 2022, since the missing value is too many, the station should be dropped

```{r}
ggplot(station, aes(x = date, y = mean_temperature_c, color = station, group = station)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  facet_grid(~station)+
  coord_flip()+
  theme_minimal() +
  labs(title = "Mean temperature",
       x = "Date",
       y = "Mean temperature (celcius degree)") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.title = element_blank(),
    legend.position = "top"
  )
```

```{r}
station <- station %>% filter(station != "Marina Barrage")
```

As the filtered table shown, some missing mean temperature value:

-   **Has minimum and maximum temperature values:** interpolate mean temperature by calculating the average of minimum and maximum temperatures

-   **Has no temperature data:** interpolate mean temperature using the `zoo` package by calculating moving averages across 7 days \>\> 15 days \>\> 30 days

For Rainfall data:

-   **Some daily total rainfall shows 0 and NA for highest rainfall:** input 0 for data with 0 total rain fall and NA for highest rainfall

-   **Daily Total Rainfall data:** apply same interpolation MA method

For wind speed data, I'll apply the same interpolation methods used for temperature.

```{r}
#| echo: false
station_filtered_c <- station %>%
  filter(is.na(mean_temperature_c))

station_filtered_c %>%
  DT::datatable(class = "display",
                caption = "Table: Missing Values in Temp_c") %>%
  formatStyle(
    columns = colnames(station_filtered_c),
    fontSize = '12px',
    fontFamily = 'Helvetica',
    lineHeight = '1.2'
  )
```

```{r}
# Step 1: Interpolate mean temperature using min and max temp
station <- station %>%
  mutate(mean_temperature_c = ifelse(
    is.na(mean_temperature_c) & !is.na(minimum_temperature_c) & !is.na(maximum_temperature_c),
    round((minimum_temperature_c + maximum_temperature_c) / 2, 2),
    mean_temperature_c
  ))

# Step 2: Define funtion for rolling mean for remaining missing values
fill_missing_with_rollapply <- function(x, window_sizes) {
  for (w in window_sizes) {
    missing_idx <- which(is.na(x))
    if (length(missing_idx) == 0) break  # Stop if no missing values remain
    
    roll_values <- rollapply(x, width = w, FUN = mean, fill = NA, align = "center", na.rm = TRUE)
    x[missing_idx] <- ifelse(is.na(x[missing_idx]), round(roll_values[missing_idx],2), x[missing_idx])
  }
  return(x)
}
# Apply rolling mean imputation with increasing window sizes
window_sizes <- c(7, 15, 30)
station$mean_temperature_c <- fill_missing_with_rollapply(station$mean_temperature_c, window_sizes)
station$minimum_temperature_c <- fill_missing_with_rollapply(station$minimum_temperature_c, window_sizes)
station$maximum_temperature_c <- fill_missing_with_rollapply(station$maximum_temperature_c, window_sizes)
station$mean_wind_speed_km_h <- fill_missing_with_rollapply(station$mean_wind_speed_km_h, window_sizes)
station$max_wind_speed_km_h <- fill_missing_with_rollapply(station$max_wind_speed_km_h, window_sizes)
station$daily_rainfall_total_mm <- fill_missing_with_rollapply(station$daily_rainfall_total_mm, window_sizes)
```

```{r}
# Input 0 for missing highest rain fall
station <- station %>%
  mutate(
    highest_30_min_rainfall_mm = ifelse(daily_rainfall_total_mm == 0 & is.na(highest_30_min_rainfall_mm), 0, highest_30_min_rainfall_mm),
    highest_60_min_rainfall_mm = ifelse(daily_rainfall_total_mm == 0 & is.na(highest_60_min_rainfall_mm), 0, highest_60_min_rainfall_mm),
    highest_120_min_rainfall_mm = ifelse(daily_rainfall_total_mm == 0 & is.na(highest_120_min_rainfall_mm), 0, highest_120_min_rainfall_mm))

# Apply MA for remaining values
station$highest_30_min_rainfall_mm <- fill_missing_with_rollapply(station$highest_30_min_rainfall_mm, window_sizes)
station$highest_60_min_rainfall_mm <- fill_missing_with_rollapply(station$highest_60_min_rainfall_mm, window_sizes)
station$highest_120_min_rainfall_mm <- fill_missing_with_rollapply(station$highest_120_min_rainfall_mm, window_sizes)


```

```{r}
kable(data.frame(Column = names(station), NAs = colSums(is.na(station))))

```

```{r}
station_filter_w <- station %>%
  filter(is.na(mean_wind_speed_km_h))
station_filter_w
```

```{r}
ggplot(station, aes(x = date, y = mean_wind_speed_km_h, color = station, group = station)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  facet_grid(~station)+
  coord_flip()+
  theme_minimal() +
  labs(title = "Mean Wind Speed",
       x = "Date",
       y = "Mean Wind Speed (km/h)") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.title = element_blank(),
    legend.position = "top"
  )
```

### 2.2.5 Save the file as RDS

```{r}
saveRDS(station, "data/station_data.rds")
```

### 2.2.5 Check time-series data

con

```{r}

ggplot(station, aes(x = date, y = daily_rainfall_total_mm, color = station, group = station)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  facet_grid(~station)+
  coord_flip()+
  theme_minimal() +
  labs(title = "Mean temperature",
       x = "Date",
       y = "Mean temperature (celcius degree)") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.title = element_blank(),
    legend.position = "top"
  )
```

```{r}
ggplot(station, aes(x = date, y = mean_temperature_c, color = station, group = station)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  facet_grid(~station)+
  coord_flip()+
  theme_minimal() +
  labs(title = "Mean temperature",
       x = "Date",
       y = "Mean temperature (celcius degree)") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.title = element_blank(),
    legend.position = "top"
  )

```

## 2.3 Generate weekly and monthly data

The group_by and summarise function will be used to aggregate weekly and monthly data:

-   Monthly data

    ```{r}
    monthly <- station %>%
      group_by(station,year,month) %>%
      summarise(
        mean_temperature = round(mean(mean_temperature_c, na.rm = TRUE), 1),
        minimum_temperature_c = min(minimum_temperature_c, na.rm = TRUE),
        maximum_temperature_c = max(maximum_temperature_c, na.rm = TRUE),
        total_rainfall = round(sum(daily_rainfall_total_mm, na.rm = TRUE),1),
        highest_30_min_rainfall_mm = max(highest_30_min_rainfall_mm, na.rm = TRUE),
        highest_60_min_rainfall_mm = max(highest_60_min_rainfall_mm, na.rm = TRUE),
        highest_120_min_rainfall_mm = max(highest_120_min_rainfall_mm, na.rm = TRUE),
        mean_wind_speed_km_h = round(mean(mean_wind_speed_km_h, na.rm=TRUE),1),
        max_wind_speed = max(max_wind_speed_km_h, na.rm = TRUE),
        .groups = "drop"
      )

    monthly <- monthly %>%
      mutate(date = ymd(paste(year, month, 1, sep = "-")),
             max_wind_speed = ifelse(max_wind_speed == -Inf, NA, max_wind_speed))
    ```

<!-- -->

-   Weekly data

    ```{r}
    weekly <- station %>%
      group_by(station,year,month,week_no) %>%
      summarise(
        mean_temperature = round(mean(mean_temperature_c, na.rm = TRUE), 1),
        minimum_temperature_c = min(minimum_temperature_c, na.rm = TRUE),
        maximum_temperature_c = max(maximum_temperature_c, na.rm = TRUE),
        total_rainfall = round(sum(daily_rainfall_total_mm, na.rm = TRUE),1),
        highest_30_min_rainfall_mm = max(highest_30_min_rainfall_mm, na.rm = TRUE),
        highest_60_min_rainfall_mm = max(highest_60_min_rainfall_mm, na.rm = TRUE),
        highest_120_min_rainfall_mm = max(highest_120_min_rainfall_mm, na.rm = TRUE),
        mean_wind_speed_km_h = round(mean(mean_wind_speed_km_h, na.rm=TRUE),1),
        max_wind_speed = max(max_wind_speed_km_h, na.rm = TRUE),
        .groups = "drop"
      )

    weekly <- weekly %>%
      mutate(date = as.Date(paste(year, week_no, 1), format = "%Y %U %u"),
             max_wind_speed = ifelse(max_wind_speed == -Inf, NA, max_wind_speed))
    ```
